### 2025-02-05 analysis/manuscript.Rmd
Describe data sources used

---
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    template: apa6.sty
    extra_dependencies: ["flafter"]
csl: apa.csl
bibliography: references.bib
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  cache.lazy = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "99%",
  fig.align = "center",
  dpi = 300,
  dev = c("png", "cairo_ps", "pdf"),
  fig.path = "fig/",
  fig.pos = "ht!", out.extra = ""
)
options(scipen = 999, digits = 2)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

```{r dependencies}
library(tidyverse, warn.conflicts = FALSE)
library(bigrquery)
library(kableExtra)
library(patchwork)
```

# Introduction

This study aims to demonstrate the suitability of open scholarly data sources for assessing the impact of transformative agreements on hybrid open access.
To achieve this, a replication study was conducted by comparing results from hoaddata, an openly available and continuously updated dataset on hybrid open access uptake based on Crossref, OpenAlex, and the cOAlition S Journal Checker Tool, with the established bibliometric databases Web of Science and Scopus.

This study focuses on the coverage of hybrid journal portfolios included in transformative agreements between 2019 and 2023.
Special attention is given to potential differences in open access uptake by country when comparing first-author affiliation data to corresponding authorships.
This is crucial because the lack of publicly available invoicing data corresponding to authorships plays an essential role in determining whether an open-access article is supported through transformative agreements.
Data on corresponding authorships have been available on the Web of Science and Scopus for much longer than in open databases such as OpenAlex, where this information is still being roled out at the time of writing.
Because of this weakness, open approaches such as hoaddata and related research use first-authorship data instead.

By conducting a large-scale comparative analysis, this study aims to

1. Determine the strengths and weaknesses of using open data sources in monitoring the impact of transformative agreements on hybrid open access publishing.
2. Assess the coverage and accuracy of open data sources compared with established bibliometric databases.
3. Evaluate the reliability of first author affiliation data as a proxy for corresponding authorship in the context of open access uptake analysis.

# Background -- Evidence base to measure the effects of transformative agreements

## Anforderungen an das Monitoring

- esac guidelines 
- gemeinsamkeiten und unterschiede zu apc (listenpreise, tatsächliche zahlungen, zentrales invoicing, rabatte, waivers) 
- insitutionen covern cas, jedoch kann es zu unterschiedlichen verrechnugnsformen führen (antielig mit förderer, splitting innerhaklbd er einrichtung)

## Bibliometrische Evidenzen

- allgmeeiner uptake 
- wachstum apcs 
- wachstum verträge (konsortien, forschung) 
- konsequenzen

# Data and methods

The aim of this study is to demonstrate the suitability of open scholarly data sources for assessing the impact of transformative agreements on hybrid open access.
To achieve this, results from hoaddata, an openly available collection of open research information regarding hybrid open access, was compared with the established bibliometric databases Web of Science and Scopus.
After describing the initial data sources used, the necessary pre-processing steps to obtain eligible articles from transformative agreements using open access evidence, author roles (first and corresponding) and affiliation data are presented.
Overall, xxxx hybrid journals from xxx agreements that published at least one open access article between 2019 and 2023. formed the basis of this study.

<!-- Replication studies, which can be found in the Life, Natural and Social Sciences, typically employ the same or comparable methods as the original study to determine whether consistent results can be obtained (https://pmc.ncbi.nlm.nih.gov/articles/PMC10019630/).

This study compares results drawn from the openly available and regularly updated hoaddata, an R package leveraging the open scholarly data sources Crossref and OpenAlex, with the in-house database of the German Competence Network for Bibliometrics (Kompetenznetzwerk Bibliometrie, KB), which comprises Web of Science and Scopus snapshots.
After describing the initial data sources used, the necessary pre-processing steps to obtain eligible articles from transformative agreements using author roles (first and corresponding) and affiliation data are presented.
Overall, xxxx hybrid journals from xxx agreements formed the basis of this study. This study covered journal articles published between 2019 and 2023.   -->

## Data sources

### hoaddata

hoaddata, developed and maintained by the author, is an R data package comprising information about the uptake of hybrid open access since 2017 from several openly available data sources.
It combines article-level metadata from Crossref and OpenAlex with transformative agreement information from the cOAlition S Journal Checker Tool (JCT), which links journal and institutional data to agreements in the ESAC registry.

More specifically, hoaddata uses Crossref, a DOI registration agency, for obtaining journal publication volume and open access status through
Creative Commons licence information relative to the published version ("version of record").
Because of limited affiliation metadata in Crossref (https://doi.org/10.31222/osf.io/smxe5), hoaddata sources  first-author affiliations from OpenAlex.
While the country in which a first-author was located was used to aggregate country-level statistics, Research Organization Registry (ROR) identifiers (ROR-ID) were used in conjunction with ESAC registry information on the duration of an agreement was used to estimate whether an article was published under active transformative agreements.
This matching benefited from the availability of the ROR-IDs in both sources.
To improve the matching, JCT data were enriched to include associated institutions, such as unviersity hospitals.

hoaddata follows good practices for computational reproducibility using R.
The package, which includes data, code, a test suite and documentation, is openly available on GitHub.
To ensure computational reproducibility while aggregating the data, a GitHub Actions continuous integration and delivery (CI/CD) workflow interfaces with the SUB Göttingen's open scholarly data warehouse based on Google BigQuery, which provides high-performant programmatic access to monthly snapshots of Crossref and OpenAlex.
The package has been regularly updated since 2022 and the version including the computation log is available on GitHub.

hoaddata is used as a data basis of the Hybrid Open Access Dashboard, a data analytics services for  library consortia and publishers to track the uptake of hybrid open access through transformative agreements.
It was also used in bibliometric research (Jahn 2025).

### Web of Science

Clarivate Analytics' Web of Science (WoS) is a well-established proprietary bibliometric database consisting of several collections [@Birkle_2020]. 
The collections considered in this study were the Science Citation Index Expanded (SCIE), the Social Sciences Citation Index (SSCI) and the Arts & Humanities Citation Index (AHCI).

The WoS provides important data points for studying spending on open access: author affiliations and roles, differentiation of journal articles into document types representing different types of journal contributions, such as original articles or reviews, and open access status information derived from OurResearch's Unpaywall, the same provider as Openalex. However, it lacks information about journals and articles under transformative agreements.

For programmatic access to article-level data, the database of the Kompetenznetzwerk Bibliometrie (KB) in Germany is used to access bibliometric data. The KB processes raw XML data provided by Clarivate Analytics, which is provided as an in-house PostgreSQL database under a uniform schema.
To support reproducibility, KB maintains annual snapshots of the database. 
Accordingly, this study used the annual snapshot from April 2024, which is considered to cover almost the entire previous publication year [@schmidt_2024_13935407].

### Scopus

Elsevier's Scopus, launched in 2004, is another widely used proprietary bibliometric database for measuring research [@Baas_2020].
Similar to the Web of Science, Scopus is selective with regard to the journals it indexes.
However, its coverage is substantially more extensive than that of the Web of Science Core collection [@Singh_2021; @Visser_2021].
With detailed metadata about article types, open access status information derived from Unpaywall, author roles, and disambiguated affiliations, Scopus also contains important data to assess open access uptake, although no direct information regarding transformative agreements was available at the time of the study.

This study used the Scopus annual snapshot of April 2024 as provided by the KB. The same KB curation effort was applied to the  Scopus raw data as for the Web of Science [@schmidt_2024_13935407].

## Data processing steps

### journal matching

### authorship records

###  Identifying eligible articles under transformative agreements

## Data records

```{r methods_overview_table, cache=TRUE}
load(here::here("data", "manuscript_data.Rda"))
methods_overview_table_df <- bind_rows(
  # Journals
  jns_ind,
  # Articles
  articles_ind,
  # OA
  oa_articles_ind,
  # Affiliation and Auhtor roles
  aff_ind
) |>
  # Human-readable indicator labels
  mutate(
    indicator = case_when(
      indicator == "ind_active_jns" ~ "Active journals",
      indicator == "ind_active_jns_core" ~ "Active journals (core)",
      indicator == "ind_active_jns_core_with_oa" ~ "Active journals (core) with OA",
      indicator == "ind_articles_total" ~ "Total published articles ",
      indicator == "ind_articles_core" ~ "Core articles",
      indicator == "ind_articles_doi" ~ "Articles with DOI",
      indicator == "ind_articles_core_doi" ~ "Core articles with DOI",
      indicator == "ind_oa_articles_all" ~ "OA articles",
      indicator == "ind_oa_articles_core" ~ "Core OA articles",
      indicator == "first_aff" ~ "First author articles",
      indicator == "cor_aff" ~ "Corresponding author articles",
    )
  )

kableExtra::kbl(
  methods_overview_table_df,
  booktabs = TRUE, 
  format.args = list(big.mark = ","), 
  digits = 1, 
  escape = TRUE, 
  position = "H",
  caption ="Coverage of hybrid journals in transformative agreements 2019-23.",
  col.names = c("", "HOAD", "Web of Science", "Scopus")
) |>
  kable_styling(position = "center") |>
  row_spec(0, bold = TRUE) |>
  pack_rows("Hybrid journal metrics", 1, 3) |>
  pack_rows("Publication metrics", 4, 5) |>
  pack_rows("Digital Object Identifier (DOI) coverage", 6, 7) |>
  pack_rows("Open Access (OA) metrics", 8, 9) |>
  pack_rows("Core articles with affiliation data", 10, 11) 
```

# results

# discussion