### 2025-03-19 analysis/cor_tables.qmd
Testing correlation

---
title: "Correlation tests"
format: 
    gfm:
      echo: true
      message: false
      warning: false
      code-fold: true
bibliography: references.bib
---

## About

The aim of this document is to determine the correlation coefficients for several hybrid open access indicators between hoaddata, Web of Science and Scopus, focusing on country affiliations between 2019 and 2023.
Related research have compared open scholarly data sources and propietory databases at the country-level using Spearman's rank correlation coefficient [@alperin2024analysissuitabilityopenalexbibliometric] and Kendal tau correlation [@Akbaritabar_2024], showing high level of correlations.

## Data preparation

```{r lood_data}
library(tidyverse)
library(corrr)
library(stats)
library(Hmisc)
library(here)

# Load data
jn_df <- readr::read_csv(here("data", "/jn_country_stats_all.csv")) |>
  filter(issn_l != "0027-8424") |>
  mutate(earliest_year = as.character(earliest_year))

# Calculate indicators by country and year
my_df_ind <- jn_df |> 
  select(-earliest_year) |> 
  group_by(country_code) |> 
  summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE))) 
```

Automate reporting

```{r automate_tables}
#' Create a correlation matrix with significance markers
#'
#' @param data A data frame containing the variables for correlation analysis
#' @param vars A character vector of variable names to include in the correlation matrix
#' @param group_name A string to be included in the title of the correlation matrix
#' @return A list with the correlation data frame and title
#' @importFrom dplyr select all_of
#' @importFrom Hmisc rcorr
#' @importFrom stringr str_to_title
create_correlation_table <- function(data, vars, group_name) {
  # Check if all variables exist in the data
  missing_vars <- setdiff(vars, names(data))
  if (length(missing_vars) > 0) {
    warning(paste("The following variables are not in the dataset:", 
                  paste(missing_vars, collapse = ", ")))
    vars <- intersect(vars, names(data))
  }

  # Select only the specified variables
  subset_data <- data %>% select(all_of(vars))

  # Calculate Spearman correlation
  corr_result <- rcorr(as.matrix(subset_data), type = "spearman")
  r <- corr_result$r
  p <- corr_result$P

  # Create a dataframe for the table
  corr_df <- as.data.frame(r)

  # Add significance markers
  for (i in 1:nrow(r)) {
    for (j in 1:ncol(r)) {
      if (!is.na(p[i, j])) {
        if (p[i, j] < 0.001) {
          corr_df[i, j] <- paste0(round(r[i, j], 2), "***")
        } else if (p[i, j] < 0.01) {
          corr_df[i, j] <- paste0(round(r[i, j], 2), "**")
        } else if (p[i, j] < 0.05) {
          corr_df[i, j] <- paste0(round(r[i, j], 2), "*")
        } else {
          corr_df[i, j] <- round(r[i, j], 2)
        }
      }
    }
  }

  # Set diagonal to "-"
  for (i in 1:nrow(corr_df)) {
    corr_df[i, i] <- "-"
  }

  return(list(df = corr_df, title = paste0(group_name, "--Spearman Correlation Matrix")))
}

#' Display correlation matrix using Pandoc table with row names
#'
#' @param corr_list A list containing the correlation data frame and title
#' @return A character string containing the Pandoc table
create_pandoc_table <- function(corr_list) {
  footnote_text <- c("\\* p < 0.05", "** p < 0.01", "*** p < 0.001")

  # Convert the correlation dataframe to include row names as a column
  corr_df <- corr_list$df
  corr_df <- rownames_to_column(corr_df, var = "Variable")

  # Add title
  table_title <- paste("###", corr_list$title, "\n\n")

  # Create the Pandoc table
  pandoc_table <- paste(capture.output(knitr::kable(corr_df, format = "pandoc")), collapse = "\n")

  # Add footnotes
  footnotes <- paste(footnote_text, collapse = "\n")

  return(paste(table_title, pandoc_table, "\n\n", footnotes, sep = "\n"))
}

```

##  Results

```{r}
#| results: asis

cor_articles <- create_correlation_table(my_df_ind, vars = c("hoad_articles", "scp_first_author_articles", "wos_first_author_articles", 
            "scp_corresponding_author_articles", "wos_corresponding_author_articles"), "Article volume")
cat(create_pandoc_table(cor_articles))
```

```{r}
#| results: asis

oa_articles <- create_correlation_table(my_df_ind, vars =  c("hoad_oa_articles", "scp_oa_first_corresponding_author_articles", "scp_oa_corresponding_author_articles",
            "wos_oa_first_author_articles", "wos_oa_corresponding_author_articles"), "Open Access volume")
cat(create_pandoc_table(oa_articles))
```

```{r}
#| results: asis
oa_prop_ind <- my_df_ind |>
  # As in Figure
  mutate(
    hoad_oa_prop = hoad_oa_articles / hoad_articles,
    scp_first_oa_prop = scp_oa_first_author_articles / scp_first_author_articles,
    scp_corresponding_oa_prop = scp_oa_corresponding_author_articles / scp_corresponding_author_articles,
    wos_first_oa_prop = wos_oa_first_author_articles / wos_first_author_articles,
    wos_corresponding_oa_prop = wos_oa_corresponding_author_articles / wos_corresponding_author_articles
  ) |>
  select(country_code, contains("prop")) |>
  create_correlation_table(vars =  c("hoad_oa_prop", "scp_first_oa_prop", 
            "scp_corresponding_oa_prop", "wos_first_oa_prop", "wos_corresponding_oa_prop"), "Open Access Uptake")
cat(create_pandoc_table(oa_prop_ind))
```

```{r}
#| results: asis
oa_prop_ind <- my_df_ind |>
  # As in Figure
  filter(!is.na(country_code), hoad_articles > 10000) |>
  mutate(
    hoad_oa_prop = hoad_oa_articles / hoad_articles,
    scp_first_oa_prop = scp_oa_first_author_articles / scp_first_author_articles,
    scp_corresponding_oa_prop = scp_oa_corresponding_author_articles / scp_corresponding_author_articles,
    wos_first_oa_prop = wos_oa_first_author_articles / wos_first_author_articles,
    wos_corresponding_oa_prop = wos_oa_corresponding_author_articles / wos_corresponding_author_articles
  ) |>
  select(country_code, contains("prop")) |>
  create_correlation_table(vars =  c("hoad_oa_prop", "scp_first_oa_prop", 
            "scp_corresponding_oa_prop", "wos_first_oa_prop", "wos_corresponding_oa_prop"), "Open Access Uptake (min 1.000 articles)")
cat(create_pandoc_table(oa_prop_ind))
```

```{r}
#| results: asis

ta_oa_articles <- create_correlation_table(my_df_ind, vars = c("hoad_ta_oa_articles", "scp_ta_oa_first_corresponding_author_articles", "scp_ta_oa_corresponding_author_articles",
            "wos_ta_oa_first_author_articles", "wos_ta_oa_corresponding_author_articles"), "Open Access volume via Transformative Agreements")
cat(create_pandoc_table(ta_oa_articles))
```

```{r}
#| results: asis
ta_oa_articles <- my_df_ind |>
    filter(!is.na(country_code), hoad_articles > 10000) |>
  create_correlation_table(vars = c("hoad_ta_oa_articles", "scp_ta_oa_first_corresponding_author_articles", "scp_ta_oa_corresponding_author_articles",
            "wos_ta_oa_first_author_articles", "wos_ta_oa_corresponding_author_articles"), "Open Access volume via Transformative Agreements (min 1,000 articles)")
cat(create_pandoc_table(ta_oa_articles))
```

```{r}
#| results: asis

ta_oa_prop <- my_df_ind |>
 #filter(hoad_ta_oa_articles > 1000) |>
  mutate(
    hoad_oa_prop = hoad_ta_oa_articles / hoad_oa_articles,
    scp_first_oa_prop = scp_ta_oa_first_author_articles / scp_oa_first_author_articles,
    scp_corresponding_oa_prop = scp_ta_oa_corresponding_author_articles / scp_oa_corresponding_author_articles,
    wos_first_oa_prop = wos_ta_oa_first_author_articles / wos_oa_first_author_articles,
    wos_corresponding_oa_prop = wos_ta_oa_corresponding_author_articles / wos_oa_corresponding_author_articles
  ) |>
  select(country_code, contains("prop")) |>
  create_correlation_table(vars =  c("hoad_oa_prop", "scp_first_oa_prop", 
            "scp_corresponding_oa_prop", "wos_first_oa_prop", "wos_corresponding_oa_prop"), "% of Hybrid OA via Agreement")
cat(create_pandoc_table(ta_oa_prop))
```

```{r}
#| results: asis

ta_oa_prop <- my_df_ind |>
 filter(hoad_ta_oa_articles > 1000) |>
  mutate(
    hoad_oa_prop = hoad_ta_oa_articles / hoad_oa_articles,
    scp_first_oa_prop = scp_ta_oa_first_author_articles / scp_oa_first_author_articles,
    scp_corresponding_oa_prop = scp_ta_oa_corresponding_author_articles / scp_oa_corresponding_author_articles,
    wos_first_oa_prop = wos_ta_oa_first_author_articles / wos_oa_first_author_articles,
    wos_corresponding_oa_prop = wos_ta_oa_corresponding_author_articles / wos_oa_corresponding_author_articles
  ) |>
  select(country_code, contains("prop")) |>
  create_correlation_table(vars =  c("hoad_oa_prop", "scp_first_oa_prop", 
            "scp_corresponding_oa_prop", "wos_first_oa_prop", "wos_corresponding_oa_prop"), "% of Hybrid OA via Agreement (min 1.000 OA articles)")
cat(create_pandoc_table(ta_oa_prop))
```

# References