### 2025-03-27 analysis/manuscript.tex

\documentclass[a4paper,man,floatsintext,longtable,noextraspace,10pt]{apa6}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{hyperref}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
% make captions italic

% number lines
% \usepackage{lineno}
% \linenumbers

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
\begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
% allow citations to break across lines
\let\@cite@ofmt\@firstofone
% avoid brackets around text for \cite:
\def\@biblabel#1{}
\def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
{\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \setlength{\leftmargin}{\cslhangindent}
  \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
{\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% tightlist
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\shorttitle{}

% flow chart
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows, positioning, calc, shapes.multipart}

% Custom styles for the flowchart
\tikzstyle{datasource} = [rectangle, rounded corners, minimum width=2.5cm, minimum height=1cm,text centered, draw=blue!60, fill=blue!5]
\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=green!60, fill=green!5]
\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=red!60, fill=red!5]
\tikzstyle{arrow} = [thick,->,>=stealth]
\tikzstyle{output} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm, text centered, draw=purple!60, fill=purple!5]

\setlength\parindent{1.27cm}

\begin{document}
\thispagestyle{otherpage}

%\maketitle

% Change \title, \date, remove \author:
\begin{large}
\textbf{Replication: Hybrid Open Access in Transformative Agreements}
\end{large}

% Add below \maketitle:
\newcommand{\orcid}{%
  \begingroup\normalfont
  \includegraphics[height=0.9em]{orcid_logo}% removed .png extension
  \endgroup
}

Najko Jahn\textsuperscript{1}\textsuperscript{*} 
(\orcid{} \href{https://orcid.org/0000-0001-5105-1463}{\color{black}{0000-0001-5105-1463}})

\textsuperscript{1} Göttingen State and University Library, University of Göttingen, Germany. \\

\textsuperscript{*} Correspondence: \href{mailto:najko.jahn@sub.uni-goettingen.de}{\color{black}{najko.jahn@sub.uni-goettingen.de}}

\section*{Abstract}
{}
{\textbf{Keywords}: hybrid open access, transformative agreements, scholarly publishing, big deals, bibliometrics}

\newpage

% QSS wants numbered sections
\setcounter{secnumdepth}{2}

\section{Introduction}\label{introduction}

This study aims to demonstrate the suitability of open scholarly data
sources for assessing the impact of transformative agreements on hybrid
open access. To achieve this, a replication study was conducted by
comparing results from hoaddata, an openly available and continuously
updated dataset on hybrid open access uptake based on Crossref,
OpenAlex, and the cOAlition S Journal Checker Tool, with the established
bibliometric databases Web of Science and Scopus.

This study focuses on the coverage of hybrid journal portfolios included
in transformative agreements between 2019 and 2023. Special attention is
given to potential differences in open access uptake by country when
comparing first-author affiliation data to corresponding authorships.
This is crucial because the lack of publicly available invoicing data
corresponding to authorships plays an essential role in determining
whether an open-access article is supported through transformative
agreements. Data on corresponding authorships have been available on the
Web of Science and Scopus for much longer than in open databases such as
OpenAlex, where this information is still being roled out at the time of
writing. Because of this weakness, open approaches such as hoaddata and
related research use first-authorship data instead.

By conducting a large-scale comparative analysis, this study aims to

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Determine the strengths and weaknesses of using open data sources in
  monitoring the impact of transformative agreements on hybrid open
  access publishing.
\item
  Assess the coverage and accuracy of open data sources compared with
  established bibliometric databases.
\item
  Evaluate the reliability of first author affiliation data as a proxy
  for corresponding authorship in the context of open access uptake
  analysis.
\end{enumerate}

\section{Background -- Evidence base to measure the effects of
transformative
agreements}\label{background-evidence-base-to-measure-the-effects-of-transformative-agreements}

\subsection{Anforderungen an das
Monitoring}\label{anforderungen-an-das-monitoring}

\begin{itemize}
\tightlist
\item
  esac guidelines
\item
  gemeinsamkeiten und unterschiede zu apc (listenpreise, tatsächliche
  zahlungen, zentrales invoicing, rabatte, waivers)
\item
  insitutionen covern cas, jedoch kann es zu unterschiedlichen
  verrechnugnsformen führen (antielig mit förderer, splitting
  innerhaklbd er einrichtung)
\end{itemize}

\subsection{Bibliometrische Evidenzen}\label{bibliometrische-evidenzen}

\begin{itemize}
\tightlist
\item
  allgmeeiner uptake
\item
  wachstum apcs
\item
  wachstum verträge (konsortien, forschung)
\item
  konsequenzen
\end{itemize}

\section{Data and methods}\label{data-and-methods}

This study investigates the suitability of open scholarly data sources
for assessing the impact of transformative agreements on hybrid open
access. As shown in Figure \ref{fig:workflow}, the methodology involved
comparing hoaddata, an openly available collection of open research
information on hybrid open access, with the bibliometric databases Web
of Science and Scopus. This section introduces the initial data sources,
followed by a presentation of the necessary data processing steps to
obtain eligible articles enabled by transformative agreements using
author roles (first and corresponding) and harmonised affiliation data.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[node distance=2cm, thick,scale=1, every node/.style={scale=0.9}]    
% Data Sources
    \node (crossref) [datasource] {Crossref};
    \node (openAlex) [datasource, right=of crossref] {OpenAlex};
    \node (jct) [datasource, right=of openAlex] {JCT};
    \node (wos) [datasource, below=of crossref] {Web of Science};
    \node (scopus) [datasource, below=of jct] {Scopus};

    % Processing Steps
    \node (hoaddata) [datasource, below=1.5cm of $(crossref)!1!(openAlex)$] {hoaddata};
    \node (kb) [process, below=1cm of $(wos)!0.5!(scopus)$] {Journal and article retrieval (including open access)};

 % Matching Process - Fixed line break syntax
    \node (matching) [process, below=1.5cm of $(hoaddata)!1!(kb)$] {Harmonising author affiliations through RoR};

 % Aggregation
    \node (transmatch) [process, below=1.5cm of $(matching)!1!(matching)$] {Estimate TA output};

    % Output
    \node (analysis) [output, below=1.5cm of $(transmatch)!1!(transmatch)$] {Aggregated analytical datasets};

    % Arrows
    \draw [arrow] (crossref) -- (hoaddata);
    \draw [arrow] (openAlex) -- (hoaddata);
    \draw [arrow] (jct) -- (hoaddata);
    \draw [arrow] (hoaddata) -- (kb);
    \draw [arrow] (wos) -- (kb);
    \draw [arrow] (scopus) -- (kb);
    \draw [arrow] (kb) -- (matching);
    \draw [arrow] (matching) -- (transmatch);
    \draw [arrow] (transmatch) -- (analysis);

    % Legend
    \node [datasource, scale=0.7] at ($(analysis)+(4,0)$) {Data Sources};
    \node [process, scale=0.7] at ($(analysis)+(4,-1)$) {Processing Steps};
    \node [output, scale=0.7] at ($(analysis)+(4,-2)$) {Output};
\end{tikzpicture}
\caption{Data processing workflow for comparing hybrid open access uptake across bibliometric data sources. The workflow shows how data from different sources (hoaddata, derived from Crossref, OpenAlex, Transformative Agreement Data dump used by the cOAlition S Journal Checker Tool (JCT), the Web of Science, and Scopus) is processed and matched using ROR IDs to enable comparative analysis. TA = Transformative Agreement.}
\label{fig:workflow}
\end{figure}

\subsection{Data sources}\label{data-sources}

\subsubsection{hoaddata}\label{hoaddata}

hoaddata, developed and maintained by the author to support open access
monitoring and research (Jahn, 2025), is an R data package that
regularly collects information on hybrid open access uptake from
multiple openly available data sources (Jahn, 2024). It combines
article-level metadata from Crossref (Hendricks et al., 2020) and
affilation metadata from OpenAlex (Priem et al., 2022) with
transformative agreement information from Transformative Agreement Data
dump used by the cOAlition S Journal Checker Tool (JCT)\footnote{\url{https://journalcheckertool.org/transformative-agreements/}},
which links journal and institutional data about participating research
organisations to agreements in the ESAC registry.

hoaddata follows good practices for computational reproducibility using
R (Marwick et al., 2018). The package, which includes data, code, a test
suite and documentation, is openly available on GitHub. To ensure
computational reproducibility while aggregating the data, a GitHub
Actions continuous integration and delivery (CI/CD) workflow handles
data retrieval from the SUB Göttingen's open scholarly data warehouse
based on Google BigQuery, which provides high-performant programmatic
access to monthly snapshots of Crossref and OpenAlex.\footnote{\url{https://subugoe.github.io/scholcomm_analytics/data.html}}
The workflow has run regularly to fetch updates from these data sources
since 2022. The package version used in this study is 0.3, containing
data from the Crossref 2024-08 dump provided to Crossref Metadata Plus
subscribers and the OpenAlex 2024-08-29 monthly dump. It covers
agreements collected between July 2021 to July 2024 from the JCT. This
version including the computation log is available on GitHub
(https://github.com/subugoe/hoaddata/releases/tag/v.0.3).

\subsubsection{Web of Science}\label{web-of-science}

Clarivate Analytics' Web of Science (WoS) is a well-established
proprietary bibliometric database consisting of several collections
(Birkle et al., 2020). Web of Science is a selective, base
research-focused database with dedicated panels to choose journals for
inclusion (Stahlschmidt \& Stephen, 2022; Visser et al., 2021). The
collections considered in this study were the Science Citation Index
Expanded (SCIE), the Social Sciences Citation Index (SSCI) and the Arts
\& Humanities Citation Index (AHCI).

These collections provides important data points for analysing open
access: author affiliations and roles, differentiation of journal
articles into document types representing different types of journal
contributions, such as original articles or reviews, and open access
status information derived from OurResearch's Unpaywall (Piwowar et al.,
2018), the same provider as OpenAlex. However, the Web of Science lacks
information about journals and articles under transformative agreements.

For programmatic access to article-level data, this study used the
database of the Kompetenznetzwerk Bibliometrie (KB) in Germany. The KB
processes raw XML data provided by Clarivate Analytics, which is
ingested into an in-house PostgreSQL database under a uniform schema. To
support reproducibility, KB maintains annual snapshots of the database.
Accordingly, this study used the annual snapshot from April 2024
(wos\_b\_202404), which is considered to cover almost the entire
previous publication year (Schmidt et al., 2024).

\subsubsection{Scopus}\label{scopus}

Elsevier's Scopus, launched in 2004, is another widely used proprietary
bibliometric database for measuring research (Baas et al., 2020).
Similar to the Web of Science, Scopus is selective with regard to the
journals it indexes. However, its journal coverage is much broader than
that of the Web of Science collections considered in this study, as it
also indexes a wider range of applied research journals (Singh et al.,
2021; Stahlschmidt \& Stephen, 2022; Visser et al., 2021). With detailed
metadata about article types, open access status information derived
from Unpaywall, author roles, and disambiguated affiliations, Scopus
also contains important data to assess open access uptake, although no
direct information regarding transformative agreements was available at
the time of the study.

This study used the Scopus annual snapshot of April 2024 as provided by
the KB (scp\_b\_202404). The same KB curation effort was applied to the
Scopus raw data as for the Web of Science (Schmidt et al., 2024).

\subsection{Data processing steps}\label{data-processing-steps}

\subsubsection{Determining hybrid journal publication
volume}\label{determining-hybrid-journal-publication-volume}

Following Jahn (2025), the starting point was a unified dataset of
several safeguarded JCT snapshots\footnote{\url{https://github.com/njahn82/jct_data}},
as provided by hoaddata. The JCT journal data in hoaddata were enriched
with ISSN variants linked to an ISSN-L. To identify hybrid journals, a
comprehensive exclusion of fully open access journals was performed
using multiple journal lists (DOAJ, OpenAlex, Bielefeld). These data
were used to determine the publication volume of hybrid journals for
each database independently.

hoaddata relies on Crossref for obtaining journal publication volume and
open access status through Creative Commons licence information relative
to the published version (``version of record''). The article metadata
included DOIs, publication dates, open access information as well as
author roles and institutional affiliations. Publication years were
determined using the earliest known date of publication in a journal. In
hoaddata, this corresponded to Crossref's issued date. For Web of
Science and Scopus, the earliest publication date was used where
available, with Scopus dates specifically determined by the KB through
version tracking of the raw data.

Many transformative agreements typically cover only certain types of
journal articles, in particular original research articles including
reviews (Borrego et al., 2021). Because of limited information on these
document types in open scholarly data (Haupka et al., 2024), hoaddata
used an extended version of Unpaywall's paratext recognition approach to
exclude non-scholarly content (Piwowar et al., 2018). To exclude
conference supplements, which are also often not covered by
transformative agreements, only articles published in regular issues,
indicated by numerical pagination, were considered. For Web of Science
and Scopus, their established, mainly accurate document type
classifications (Donner, 2017; Maisano et al., 2025) were used to
identify original research articles and reviews, referred to as original
articles throughout this study.

\subsubsection{Identifying open access articles in hybrid
journals}\label{identifying-open-access-articles-in-hybrid-journals}

Articles in hybrid journals were considered open access when they were
made freely available under a Creative Commons license on publishers'
platforms. While hoaddata sourced this information from Crossref license
metadata, Web of Science and Scopus relied on Unpaywall as evidence
source. Unpaywall also uses Crossref license metadata, but supplements
them by parsing publisher websites directly, addressing cases where
publishers do not provide machine-readable Creative Commons license
information (Piwowar et al., 2018). This additional parsing remains
necessary despite transformative agreement workflows recommend the
depositing license information during DOI registration (Geschuhn \&
Stone, 2017). Both Web of Science and Scopus defined hybrid open access
consistently as content available under Creative Commons licenses on
publisher platforms according to their
documentations\footnote{\url{https://webofscience.help.clarivate.com/en-us/Content/open-access.html}}\footnote{\url{https://blog.scopus.com/posts/scopus-filters-for-open-access-type-and-green-oa-full-text-access-option}},
distinguishing it from bronze open access that lack such explicit
license information, or use publisher-specific licenses.

\subsubsection{Harmonising author affiliations across
databases}\label{harmonising-author-affiliations-across-databases}

Author affiliations were retrieved for both first and, if available,
corresponding authors to prepare the linking between articles and
institutions covered by transformative agreements. To improve the data
retrieval, JCT institution data was enriched with ROR-IDs from
associated institutions, such as university hospitals or institutes of
large research organisations such as the Max Planck Society, according
to OpenAlex' institution entity. To handle different address variants,
database-specific affiliation identifiers were used: ROR-IDs from
OpenAlex for hoaddata, affiliation enhanced names for Web of Science,
and Scopus Affiliation Identifier. Additionally, ISO country codes were
retrieved for each author's address to compile country-level statistics.
These country codes for Web of Science and Scopus were provided by the
KB.

Because neither Web of Science nor Scopus support ROR IDs, the
institution identifier used by the JCT, a two-step matching process was
implemented to harmonise affiliation data. First, 2,782,540 articles
from 6,457 institutions with ROR-IDs in the JCT data since 2017
(according to hoaddata) were processed to map first authors' ROR-IDs to
corresponding proprietary affiliation identifier in Web of Science and
Scopus using DOI matching. Then, an algorithm selected the most frequent
ROR ID and proprietary identifier pairs to handle multiple affiliations
and organisational hierarchy differences.

This process linked 6,375 ROR IDs to 4,894 Scopus Affiliation IDs, and
6,034 ROR IDs to 2,422 enhanced affiliation strings in the Web of
Science. Quality evaluation through random sampling of 50 pairs revealed
an error rate of 22\% for Web of Science (11 mismatches) and 6\% for
Scopus (3 mismatches). Upon inspection, these mismatches primarily
occurred with less-represented institutions having only a few
publications, introduced through multiple affiliations of single
authors. The difference between databases suggests that Scopus's
affiliation control aligns more closely with ROR than that of the Web of
Science.

\subsubsection{Estimating open access in hybrid journals covered by
transformative
agreements}\label{estimating-open-access-in-hybrid-journals-covered-by-transformative-agreements}

Based on these generated matching tables, articles eligible under
transformative agreements could also be retrieved from Web of Science
and Scopus, although they did not contain the ROR IDs used by the JCT.
The estimation of eligible articles followed Jahn (2025) and included a
matching of both journals and participating institutions. The matching
also took into account the duration of agreements according to the ESAC
registry, with only those matches where an agreement was actually in
place being considered for subsequent analysis. A related study (Jonge
et al., 2025), applied to publications funded by the Dutch Research
Council (NWO) and validated against internal invoicing data, confirmed
that such matching can accurately identify most articles under
transformative agreements.

\subsection{Data records}\label{data-records}

As a result of the comprehensive data processing described above,
datasets on open access in hybrid journals included in transformative
agreements were aggregated for each database at country and journal
level by year. Table~\ref{tbl-methods_overview_tab} provides a general
overview of the coverage between 2019 and 2023. It shows that the
majority of hybrid journals published at least one original research
article or review marked as original during the five-year period. These
journals formed the basis for the subsequent calculation of
article-level indicators.

\begin{table}

\caption{\label{tbl-methods_overview_tab}Coverage of hybrid journals in
transformative agreements 2019-23.}

\centering{

\centering
\begin{tabular}[t]{lrrr}
\toprule
\textbf{} & \textbf{hoaddata*} & \textbf{Web of Science} & \textbf{Scopus}\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Hybrid journal coverage}}\\
\hspace{1em}Active journals & 12,890 & 8,655 & 11,888\\
\hspace{1em}Active journals (core) & 12,888 & 8,655 & 11,878\\
\hspace{1em}Active journals (core) with OA & 11,348 & 8,392 & 11,313\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Publication volume}}\\
\hspace{1em}Total published articles & 9,740,015 & 8,616,053 & 8,117,644\\
\hspace{1em}Original articles & 8,158,425 & 6,708,083 & 7,317,703\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Digital Object Identifier (DOI) coverage}}\\
\hspace{1em}Articles with DOI & 9,740,015 & 7,713,796 & 8,105,112\\
\hspace{1em}Original articles with DOI & 8,158,425 & 6,695,661 & 7,314,327\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Open Access (OA) metrics}}\\
\hspace{1em}OA articles & 998,699 & 1,112,758 & 974,099\\
\hspace{1em}Original OA articles & 969,817 & 1,019,784 & 922,578\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Original articles with affiliation data}}\\
\hspace{1em}First author articles & 7,242,542 & 6,294,855 & 7,232,017\\
\hspace{1em}Corresponding author articles & 5,534,207 & 6,291,441 & 6,898,487\\
\bottomrule
\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{*} Journal article metadata from Crossref, except affiliations from OpenAlex}\\
\end{tabular}

}

\end{table}%

While hoaddata only covers articles with a DOI, Scopus and Web of
Science publication indicators were calculated using their database
identifier. A subsequent comparison of DOI coverage shows that
non-original articles in Web of Science often lacked a DOI. This was
particularly the case for meeting abstracts, which are notably prevalent
in Health Sciences journals (Melero-Fuentes et al., 2025) and are not
indexed by Scopus (Donner, 2017). Open access indicators were aggregated
by DOI, as Unpaywall only collects information on open access status for
articles with a DOI. A closer look at original articles with affiliation
data, and in line with related research (Zhang et al., 2024), reveals a
lack of affiliation data in the case of OpenAlex, the affiliation data
source used by hoaddata, compared to Web of Science and Scopus. In
particular, only about two-thirds of the articles examined provided
corresponding author affiliations. For first authors, the proportion was
89\%. At the time of writing, OpenAlex disclosed limited coverage of
corresponding authorhip data \footnote{\url{https://docs.openalex.org/api-entities/works/work-object/authorship-object\#is_corresponding}}.
Therefore, only first author data for hoaddata were considered in the
following analysis.

\subsection{Data analysis}\label{data-analysis}

\section{Results}\label{results}

This section first presents the indexing coverage of hybrid open access
by comparing the data from open data sources with the proprietary
bibliometric databases Scopus and Web of Science. Then, using the same
methods, indicators at the publisher and country level are calculated
independently for each database and compared with each other to assess
the suitability of the bibliometric databases for investigating
transformative agreements.

\subsection{Coverage comparision}\label{coverage-comparision}

\subsubsection{Overview}\label{overview}

Figure \ref{fig-upset_coverage_results} presents the coverage of hybrid
journals included in transformative agreements, visualising the
intersections of journals and articles across the examined databases as
an UpSet graph (Krassowski, 2020; Lex et al., 2014). The analysis
included hybrid journals that published at least one open access article
between 2019 and 2023, based on open access status information from each
database. Only original research articles and reviews were considered in
the analysis.

\begin{figure}[ht!]

\centering{

\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/fig-upset_coverage_results-1.pdf}

}

\caption{\label{fig-upset_coverage_results}Upset graph}

\end{figure}%

Journal coverage analysis revealed that 66\% (n = 7,970) of hybrid
journals included in transformative agreements were indexed in all three
databases (Figure \ref{fig-upset_coverage_results}A). The second-largest
set consisted of journals indexed in both hoaddata and Scopus,
comprising 21\% (n = 2,595) of hybrid journals. Notably, 6\% (n = 739)
of journals were exclusively contained in hoaddata, while another 6\% (n
= 748) were only found in the proprietory database Scopus. Of these, 354
were also available in the Web of Science. Upon inspection, this group
of hybrid journals exclusively covered in the proprietary data sources
mainly represented hybrid journals for which no open access evidence
could be retrieved from Crossref, the open access evidence source for
hoaddata.

In terms of article coverage, Figure \ref{fig-upset_coverage_results}B
shows the total publication volume per combination in terms of DOI
availability. The largest set of hybrid journals, which includes all
three data sources, also contains the largest number of articles. In
total, these journals recorded 6,289,687 overlapping articles,
represented by the blue bar. They represented 94\% of original articles
with DOI indexed in the Web of Science sample, and 86\% in Scopus.
Another 657,697 articles were exclusive to both Scopus and hoaddata.
Exclusively in hoaddata were 177,110 articles, and exclusively in the
proprietary databases were 325,194 articles.

Figure \ref{fig-upset_coverage_results}B also shows the surplus of
articles with DOI that were only available via hoaddata (grey area). In
case of hybrid journals covered by all three data sources, 1,023,882
DOIs were only present in hoaddata. After validation at the DOI level
using the KB databases and manual inspection, the main reasons for
missing DOI coverage in the proprietary database were insufficient
classification of journal content as original research articles and
reviews during the compilation of hoaddata. Particularly, letters and
editorials could not be fully detected. Paratext recognition failed in
37\% of DOIs to identify non-scholarly content such as front matters or
reviewer lists, which are generally not indexed by Scopus and Web of
Science. To a lesser extent, differences in publication and indexing
dates were a reason for non-overlapping DOIs.

Using overlapping DOIs, the publication volume between 2019 and 2023 was
also calculated for each journal. Figure
\ref{fig-upset_coverage_results}C illustrates the distribution for each
combination. It shows that there is a large spread across the journals
covered by all three data sources. Furthermore, journals in this set
published more on average than journals covered by only one or two data
sources. In particular, the journals covered exclusively by hoaddata
were substantially smaller than those covered by all three sources. Upon
inspection, these were often newly launched hybrid journals, which
explains the relatively low five-year publication volume. An example is
\emph{Digital society} that published 86 articles. This hybrid journal
was launched in 2022, being covered by various Springer Nature
transformative agreements since then.

\subsubsection{Coverage by publisher
portfolio}\label{coverage-by-publisher-portfolio}

Figure \ref{fig-upset_coverage_results_publisher} presents the coverage
of hybrid journals in transformative agreements across data sources from
2019 to 2023 with a focus on publisher portfolios. The analysis
highlights the dominance of the three largest publishers--Elsevier,
Springer Nature, and Wiley--, which collectively accounted for 47\% of
hybrid journals and 62\% of articles published during this five-year
period. In terms of article volume, Elsevier led with 2,441,358 articles
(33\% of the total) published across 1,951 hybrid journals (16\% of the
total). Springer Nature followed with 1,247,578 articles (17\%) in
2,311, although recording the largest number of hybrid journals (19\%).
Wiley accounted for 858,939 articles (12\%) in 1,382 hybrid journals
(11\%). The remaining 54 publishers collectively accounted for 2,749,847
articles (38\%) in 6,476 hybrid journals (53\%).

\begin{figure}[ht!]

\centering{

\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/fig-upset_coverage_results_publisher-1.pdf}

}

\caption{\label{fig-upset_coverage_results_publisher}Upset graph
publisher}

\end{figure}%

The three largest publishers--Elsevier, Springer Nature and Wiley--were
best represented in the exclusive intersection of all three data sources
(hoaddata, Scopus, and Web of Science). Together, they comprised 4,384
hybrid journals (55\% of the intersectional set) and dominated article
coverage (n = 4,174,315; 66\%), as determined through shared DOIs. When
examining publication volume per journal, (Figure
\ref{fig-upset_coverage_results_publisher}C), Elsevier published, on
average, the largest journals, followed by Springer Nature and Wiley.

Comparing publisher portfolios across different indexing sets
demonstrates that publisher were not represented uniformly. Notably,
Springer Nature exhibited 519 hybrid journals exclusively indexed in
both hoaddata and Scopus. This set included journals from the Chinese
Academy of Science, German-language medical journals, and Eastern
European publications including the \emph{Journal of Mathematical
Sciences}, which also publishes English-language translations of
Russian-language works. Additionally, this subset included titles with a
more broader disciplinary focus such as \emph{SN Computer Science} and
newly launched hybrid journals like \emph{Nature Computational Science},
which started in 2021 and was indexed in Scopus but not yet in Web of
Science. The set also captured ceased journals, providing further
insights into the dynamics of journal publishing.

Examining publisher portfolios not covered by hoaddata but present in
Scopus or Web of Science identified several publishers with missing CC
license information in Crossref. In particular, Emerald represented 322
journals with 86,409 articles, AIP Publishing accounted for 24 journals
with 64,898 articles, and World Scientific recorded 87 journals and
42,531 articles. In total, 9 publishers were not represented in
Crossref.

An inspection of individual journals also uncovered discrepancies in
Unpaywall's open access identification for certain publishers that
typically deposit CC licenses with Crossref. Notably, some
subscription-only journals contained one or two articles erroneously
tagged as hybrid open access by Unpaywall, which were subsequently
reflected in Scopus and Web of Science. Examples of such
misclassifications include Elsevier's \emph{Journal of Bioscience and
Bioengineering} and Springer Nature's \emph{Journal of Mechanical
Science and Technology}.

\subsection{Open Access Indicator
Comparision}\label{open-access-indicator-comparision}

This section examines the uptake of open access in hybrid journals,
focusing on the influence of transformative agreements across hoaddata,
Scopus, and Web of Science. The aim was to assess whether consistent
results can be derived from these data sources despite differences in
coverage and methodologies. Following Jahn (2025), indicators were
calculated for each data source and comprise the number and proportion
of open access articles, including those enabled by transformative
agreements, from 2019 to 2023. For Web of Science and Scopus, the impact
of transformative agreements was estimated using both first and
corresponding authorships, while hoaddata indicator calculation was
limited to first author affiliations.

\begin{figure}[ht!]

\centering{

\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/fig-uptake_overview-1.pdf}

}

\caption{\label{fig-uptake_overview}Open Access indicato}

\end{figure}%

\subsubsection{Overview}\label{overview-1}

Figure \ref{fig-uptake_overview}A shows a moderate growth of open access
in hybrid journals, which is consistent across hoaddata (black line),
Scopus (orange line), and Web of Science (blue line). According to
hoaddata, hybrid open access uptake increased from \% (n = ) in 2019 to
\% (n = ) in 2023. Similarly, Scopus recorded an growth from 6.5\% (n =
84,648) in 2019 to 19\% (n = 322,850) in 2023. However, Web of Science
recorded higher open access uptake in early years, before converging in
2023, from 11\% (n = 137,202) in 2019 to 20\% (n = 255,481) in 2023,
suggesting an different approach towards labeling hybrid open access by
the Web of Science.

Similarly, hybrid open access by transformative agreements increased
between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and C). Trends
were consistent for first (dashed line) and corresponding author (dotted
line) affiliations. According to Scopus, 479,297 open access articles
could be attributed to transformative agreements based on first author
metadata (increasing from 15,341 to 206,084) and 489,262 using
corresponding author metadata (from 15,444 to 210,816). Web of Science
recorded 493,028 open access articles via transformative agreements
using first author metadata (increasing from 21,871 to 174,126) and
500,076 using corresponding author metadata (from 22,092 to 177,030).
hoaddata, lacking corresponding author data, linked 0 articles to
transformative agreements but showed slower growth than Scopus and Web
of Science, from 0 in 2019 to 0 in 2023.

From 2021 (hoadata, Scopus) resp. 2022 (Web of Science), transformative
agreements enabled the majority of hybrid open access. For first
authors, the share ranged between \% (hoaddata), 64\% (Scopus), and 68
\% (Web of Science) in 2023. For corresponding authors, the shares were
slightly larger, with Scopus recording 65\% and Web of Science 69\% in
2023. However, substantial hybrid open access was still facilitated
outside transformative agreements, likely through APCs paid from
discretionary research funds (Jahn et al., 2022; Suber, 2012).

\subsubsection{Open access by
publishers}\label{open-access-by-publishers}

\begin{figure}[ht!]

\centering{

\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/fig-uptake_publisher-1.pdf}

}

\caption{\label{fig-uptake_publisher}Open Access indicato}

\end{figure}%

When considering open access trends by publisher (see Figure
\ref{fig-uptake_publisher}), the observed differences in early uptake
rates between hoaddata and Scopus compared to Web of Science can be
largely attributed to articles published in Elsevier hybrid journals,
the largest publisher in our sample. Both hoadata and Scopus reported an
steady increase in open access uptake between 2019 and 2023 (hoaddata
from 4\% to 13\%; Scopus from 5\% to 15\%). In contrast, Elsevier's
share remained relatively constant, increasing only slightly from 14\%
to 15\% according to the Web of Science. Upon inspection, this
discrepancy primarily stemmed from articles in the publisher's open
archive. These articles, made freely available after an embargo period
under Elsevier's user license, were tagged as hybrid open access in Web
of Science, even though its documentation\footnote{\url{https://webofscience.help.clarivate.com/en-us/Content/open-access.html}}
specified that only articles under a CC license variant were considered.
Previous research (Haustein et al., 2024; Jahn et al., 2022) has shown
that Elsevier provided a substantial portion of its articles under this
license, explaining the relatively large and stable share of open access
over the years.

Differences in open access evidence are also evident for Wiley.
Specifically, Web of Science and Scopus recorded a drop in 2021 and 2022
compared to hoaddata. For these two years, hoaddata reported 35,308 more
open access articles than Scopus and 32,491 more than Web of Science.
This discrepancy is presumably because of challenges in fetching
full-texts by Unpaywall, the open access evidence source for Scopus and
Web of Science. According to Unpaywall's software version history,
Wiley's publisher platform redirects prevented Unpaywall from parsing
license information from full-texts.\footnote{See Unpaywall version
  history related to Wiley fixes:
  \url{https://github.com/search?q=repo\%3Aourresearch\%2Foadoi+wiley&type=commits}}
hoaddata, which relies solely on Crossref metadata for open access
identification, was unaffected by these issues.

Despite these differences in open access evidence, the three data
sources show consistent temporal trends in hybrid open access enabled by
transformative agreements (see Figure \ref{fig-uptake_publisher}B and
C). Wiley emerged as the fastest-growing publisher in terms of open
access uptake, with more than 30\% of articles in hybrid journals
reported as open access in 2023 across the examined data sources,
followed by Springer Nature and Elsevier, recording a comparable late
uptake, which is consistent with the publisher's historical reluctance
to engage in negotations with library consortia. However, by 2023, the
share of open access enabled by transformative agreements appeared to
stabilise for all three publishers (see Figure
\ref{fig-uptake_publisher}C). Interestingly, the differences between
first and corresponding author affiliations were more pronounced at the
publisher level. In Scopus, for example, the share of open access via
transformative agreements measured by corresponding authorship was
greater for Springer Nature in 2023 than when using first authorship.

\subsubsection{Open access by country}\label{open-access-by-country}

When comparing countries, consistent patterns were observed across data
sources for the five-year period 2019 to 2023. Figure
\ref{fig-uptake_country} presents hybrid open access indicators by
country, comparing hoaddata (x-axis) with Web of Science and Scopus
(y-axis). Indicators calculated from these proprietary databases are
shown for both first and corresponding authors, with full counting used
to account for multiple country affiliations.

\begin{figure}[ht!]

\centering{

\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/fig-uptake_country-1.pdf}

}

\caption{\label{fig-uptake_country}Open Access by country}

\end{figure}%

In terms of article output by country (see Figure
\ref{fig-uptake_country}A), strong positive correlation was observed
across the data sources and author roles (Spearman rank correlation
\(\rho > .9\), \(p < 0.001\)). Between 2019 and 2023, China was the most
productive country, followed by the United States and, by a certain
margin, India, the United Kingdom, and Germany. Analysis of authorship
roles revealed minimal variation, indicating that first and
corresponding authors were typically from the same country.

When examining the percentage of open access articles in hybrid journals
(see Figure \ref{fig-uptake_country}B), a different pattern emerged.
Authors affiliated with institutions from medium-sized European
countries, such as Sweden, the Netherlands, Finland, and Hungary,
provided a large proportion of their articles in open access. Germany
and the United Kingdom also had approximately 40\% of their output
available as open access. In contrast, non-OECD countries showed notably
lower adoption of hybrid open access, with South Africa being the only
BRICS member well-represented in the data. The United States also
demonstrated a relatively low proportion of open access articles. These
findings were consistent across all databases. However, France was
better represented in Web of Science, likely due to its agreement with
Elsevier starting in 2019, which allowed delayed open access under the
publisher's user license (Rabesandratana, 2019). This licence was not
classified as hybrid open access in either Scopus or hoaddata. In all
cases, Spearman rank correlations were \(\rho > .9\), \(p < 0.001\),
showing a high level of correlations between the databases and
authorship roles considered.

Transformative agreements appeared to be a key driver of national open
access growth (see Figure \ref{fig-uptake_country}C and D). OECD members
accounted for the majority of open access articles enabled by
transformative agreements. As a notable exception, South Africa also
featured prominently, as the South African National Library and
Information Consortium (SANLiC) successfully negotiated transformative
agreements with major publishers from 2022 onward.\footnote{\url{https://sanlic.ac.za/read-and-publish-agreements/}}
Results were consistent across data sources. However, Wiley's open
access surplus in hoaddata led to better rankings for countries where
Wiley played a substantial role, such as Germany, where the DEAL
consortium negotiated its first transformative agreements with Wiley
that started in July 2019.

A closer examination of the proportion of hybrid open access enabled by
transformative agreements (see Figure \ref{fig-uptake_country}D)
revealed that the French Elsevier agreement on delayed open access was
not included in JCT data, as it did not qualify as a transformative
agreement. Medium-sized European countries again showed a high
proportion of hybrid open access through transformative agreements,
highlighting the impact of this licensing model across all three data
sources. In contrast, the United States had a low proportion of hybrid
open access enabled by agreements, suggesting that a substantial number
of open access articles were likely financed through other means, such
as article processing charges (APCs).

In all cases, strong positive correlations were observed using
Spearman's rank correlation: \(\rho > .9\), \(p < 0.001\) between data
sources and authorship roles, when considering countries with an minimum
of 1,000 open access articles enabled by transformative agreements
between 2019 and 2023. When this limitation was removed, the correlation
remained strong but slightly lower (\(\rho > .85\), \(p < 0.001\)). This
difference may signal countries where only a few institutions had
transformative agreements in place, as opposed to those participating in
national consortia with broader participation.

\section{Discussion}\label{discussion}

Investigating over 13,000 hybrid journals included in the cOAlition
Journal Checker Tool shows a substantial increase in open access enabled
by transformative agreements between 2019 and 2023, although most
articles in these journals remained paywalled during this five-year
period. While transformative agreements accounted for the majority of
open access, a considerable number of open access articles are likely to
have resulted from the payment of individual publication fees. The
results confirm that transformative agreements and hybrid open access
continue to be concentrated among a few large commercial publishers.
They also highlight that high-income European countries in particular
are using transformative agreements to make most of their article output
open access, although South Africa showed similar levels of open access
adoption. These trends were consistent across the investigated open data
source hoaddata, derived from Crossref and OpenAlex, and the established
proprietary bibliometric databases Scopus and Web of Science.

Comparing the data sources examined, the study results show strong
correlations by country affiliation, despite differences in journal and
article coverage, and in the availability of metadata on corresponding
authorship and open access status. However, several notable
discrepancies warrant discussion. They stem not only from known
limitations of the open scholarly metadata sources Crossref and
OpenAlex, but also from the proprietary databases Scopus and Web of
Science.

The coverage analysis reveals that hybrid journals are well indexed in
all three data sources, particularly in terms of article coverage.
Differences can be found for journals targeting practitioners or local
non-English language communties, with many such titles indexed
exclusively in Crossref and Scopus. Using the open index Crossref
demonstrated particular strength in identifying newly established hybrid
journals, a notable finding given that transformative agreements
primarily target existing subscription-based journals.

The situation of hybrid journal publishing is therefore different from
that of fully open access journals. Comparing the coverage of OpenAlex,
Scopus and Web of Science, Simard et al. (2024) indicate that only half
of the fully open access journals listed in the DOAJ are also indexed in
Scopus and Web of Science. Notably, journals that charge no publication
fees (``diamond journals'') are absent from the selective Web of
Science, which reinforces existing disparities in the indexing of
underrepresented research communities and regions in selective
bibliometric databases (ref).

A notable advantage of using Scopus and Web of Science to study the
impact of transformative agreements is their comprehensive
differentiation of journal articles by document type, as many agreements
restrict funding to eligible article types. This resulted in slightly
higher rates of open access uptake compared to hoaddata.

A similar limitation of studying open access with open scholarly data
sources is the often reported lack of corresponding authorship
information. However, indicators using first authors, which have often
been used as a proxy for determining open access funding, and
corresponding authors show a high level of correlation, which is not
surprising given disciplinary norms in scholarly publishing. First
authors are often considered to have carried out the main research
underlying a paper, while the corresponding author is often recognised
as having supervised a paper. It is therefore not surprising that both
are strongly correlated, suggesting that first and corresponding authors
share the same country affiliation. Moreover, in most cases the first
author is indeed the same as the corresponding author. As such, first
authorship can be used confidentially in large-scale studies and those
focusing on national consortia to assess the impact of transformative
agreements.

However, affiliation data is sparser in OpenAlex than in Scopus and Web
of Science, resulting in slightly more articles that could be attributed
to transformative agreements. However, in order to establish a link, a
thorough mapping of the database affiliation identifier to the ROR ID,
the JCT's identifier for institutional participation in transformative
agreements, is required. In contrast, ROR IDs are the canonical
organisation identifier in OpenAlex, allowing for less ambiguous
matching between agreement and publication.

Comparison of scholarly databases is also valuable in identifying
differences in open access identification. Not all publishers deposit a
Creative Commons licence with Crossref, the citation authority. A
collaboration with Scopus and Web of Science helped to identify such
publishers, as their underlying unpaywall data also parses licence
information from websites. However, there were significant differences
between Scopus and Web of Science in terms of open access indexing,
particularly with regard to the delayed open access provided by
Elsevier. In the case of Wiley, however, there were also problems with
the analysis of the Unpaywall. This suggests that hybrid open access
studies should not only refer to a single source of evidence. Instead,
open access tags can be validated by the combiantion of seevral data
sources, as in our case where metadata from publishers was compared with
data from open access discovery services such as Unpaywall. It is also
particularly useful to include a curated list of journals under the
hybrid model to avoid misclassifications based on a journal's business
model (Visser et al., 2021).

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Baas_2020}
Baas, J., Schotten, M., Plume, A., Côté, G., \& Karimi, R. (2020).
Scopus as a curated, high-quality bibliometric data source for academic
research in quantitative science studies. \emph{Quantitative Science
Studies}, \emph{1}(1), 377--386.
\url{https://doi.org/10.1162/qss_a_00019}

\bibitem[\citeproctext]{ref-Birkle_2020}
Birkle, C., Pendlebury, D. A., Schnell, J., \& Adams, J. (2020). Web of
science as a data source for research on scientific and scholarly
activity. \emph{Quantitative Science Studies}, \emph{1}(1), 363--376.
\url{https://doi.org/10.1162/qss_a_00018}

\bibitem[\citeproctext]{ref-Borrego_2020}
Borrego, Á., Anglada, L., \& Abadal, E. (2021). Transformative
agreements: Do they pave the way to open access? \emph{Learned
Publishing}, \emph{34}(2), 216--232.
\url{https://doi.org/10.1002/leap.1347}

\bibitem[\citeproctext]{ref-Donner_2017}
Donner, P. (2017). Document type assignment accuracy in the journal
citation index data of {Web of Science}. \emph{Scientometrics},
\emph{113}(1), 219--236. \url{https://doi.org/10.1007/s11192-017-2483-y}

\bibitem[\citeproctext]{ref-Geschuhn_2017}
Geschuhn, K., \& Stone, G. (2017). It's the workflows, stupid! What is
required to make {``offsetting''} work for the open access transition.
\emph{Insights the {UKSG} Journal}, \emph{30}(3), 103--114.
\url{https://doi.org/10.1629/uksg.391}

\bibitem[\citeproctext]{ref-Haupka_2024}
Haupka, N., Culbert, J. H., Schniedermann, A., Jahn, N., \& Mayr, P.
(2024). \emph{Analysis of the publication and document types in
OpenAlex, web of science, scopus, pubmed and semantic scholar}. arXiv.
\url{https://doi.org/10.48550/ARXIV.2406.15154}

\bibitem[\citeproctext]{ref-Haustein_APC_2024}
Haustein, S., Schares, E., Alperin, J. P., Hare, M., Butler, L.-A., \&
Schönfelder, N. (2024). \emph{Estimating global article processing
charges paid to six publishers for open access between 2019 and 2023}.
\url{https://arxiv.org/abs/2407.16551}

\bibitem[\citeproctext]{ref-Hendricks_2020}
Hendricks, G., Tkaczyk, D., Lin, J., \& Feeney, P. (2020). Crossref: The
sustainable source of community-owned scholarly metadata.
\emph{Quantitative Science Studies}, \emph{1}(1), 414--427.
\url{https://doi.org/10.1162/qss_a_00022}

\bibitem[\citeproctext]{ref-hoaddata}
Jahn, N. (2024). \emph{Hoaddata: Data about hybrid open access journal
publishing}. \url{https://github.com/subugoe/hoaddata}

\bibitem[\citeproctext]{ref-Jahn_2025}
Jahn, N. (2025). How open are hybrid journals included in transformative
agreements? \emph{Quantitative Science Studies}, 1--39.
\url{https://doi.org/10.1162/qss_a_00348}

\bibitem[\citeproctext]{ref-Jahn_2021}
Jahn, N., Matthias, L., \& Laakso, M. (2022). Toward transparency of
hybrid open access through publisher-provided metadata: An article-level
study of {Elsevier}. \emph{Journal of the Association for Information
Science and Technology}, \emph{73}(1), 104--118.
\url{https://doi.org/10.1002/asi.24549}

\bibitem[\citeproctext]{ref-de_Jonge_2025}
Jonge, H. de, Kramer, B., \& Sondervan, J. (2025). \emph{Tracking
transformative agreements through open metadata: Method and validation
using dutch research council NWO funded papers.}
\url{https://doi.org/10.31222/osf.io/tz6be_v1}

\bibitem[\citeproctext]{ref-ComplexUpset}
Krassowski, M. (2020). \emph{ComplexUpset}.
\url{https://doi.org/10.5281/zenodo.3700590}

\bibitem[\citeproctext]{ref-Lex_2014}
Lex, A., Gehlenborg, N., Strobelt, H., Vuillemot, R., \& Pfister, H.
(2014). UpSet: Visualization of intersecting sets. \emph{IEEE
Transactions on Visualization and Computer Graphics}, \emph{20}(12),
1983--1992. \url{https://doi.org/10.1109/tvcg.2014.2346248}

\bibitem[\citeproctext]{ref-Maisano_2025}
Maisano, D. A., Mastrogiacomo, L., Ferrara, L., \& Franceschini, F.
(2025). A large-scale semi-automated approach for assessing
document-type classification errors in bibliometric databases.
\emph{Scientometrics}. \url{https://doi.org/10.1007/s11192-025-05244-y}

\bibitem[\citeproctext]{ref-Marwick_2018}
Marwick, B., Boettiger, C., \& Mullen, L. (2018). Packaging data
analytical work reproducibly using {R} (and friends). \emph{The American
Statistician}, \emph{72}(1), 80--88.
\url{https://doi.org/10.1080/00031305.2017.1375986}

\bibitem[\citeproctext]{ref-Melero_Fuentes_2025}
Melero-Fuentes, D., Aguilar-Moya, R., Valderrama-Zurián, J.-C., \&
Gorraiz, J. (2025). Evolution and effect of meeting abstracts in JCR
journals. \emph{Journal of Informetrics}, \emph{19}(1), 101631.
\url{https://doi.org/10.1016/j.joi.2024.101631}

\bibitem[\citeproctext]{ref-Piwowar_2018}
Piwowar, H., Priem, J., Larivière, V., Alperin, J. P., Matthias, L.,
Norlander, B., Farley, A., West, J., \& Haustein, S. (2018). The state
of {OA}: A large-scale analysis of the prevalence and impact of open
access articles. \emph{{PeerJ}}, \emph{6}, e4375.
\url{https://doi.org/10.7717/peerj.4375}

\bibitem[\citeproctext]{ref-priem2022openalex}
Priem, J., Piwowar, H., \& Orr, R. (2022). \emph{OpenAlex: A fully-open
index of scholarly works, authors, venues, institutions, and concepts}.
\url{https://arxiv.org/abs/2205.01833}

\bibitem[\citeproctext]{ref-Rabesandratana_2019}
Rabesandratana, T. (2019). Elsevier deal with france disappoints
open-access advocates. \emph{Science}.
\url{https://doi.org/10.1126/science.aba5656}

\bibitem[\citeproctext]{ref-schmidt_2024_13935407}
Schmidt, M., Rimmert, C., Stephen, D., Lenke, C., Donner, P., Gärtner,
S., Taubert, N., Bausenwein, T., \& Stahlschmidt, S. (2024). \emph{The
data infrastructure of the {German Kompetenznetzwerk Bibliometrie}: An
enabling intermediary between raw data and analysis}. Zenodo.
\url{https://doi.org/10.5281/zenodo.13935407}

\bibitem[\citeproctext]{ref-Simard_2024}
Simard, M.-A., Basson, I., Hare, M., Lariviere, V., \& Mongeon, P.
(2024). \emph{The open access coverage of OpenAlex, scopus and web of
science}. \url{https://doi.org/10.48550/ARXIV.2404.01985}

\bibitem[\citeproctext]{ref-Singh_2021}
Singh, V. K., Singh, P., Karmakar, M., Leta, J., \& Mayr, P. (2021). The
journal coverage of web of science, scopus and dimensions: A comparative
analysis. \emph{Scientometrics}, \emph{126}(6), 5113--5142.
\url{https://doi.org/10.1007/s11192-021-03948-5}

\bibitem[\citeproctext]{ref-Stahlschmidt_2022}
Stahlschmidt, S., \& Stephen, D. (2022). From indexation policies
through citation networks to normalized citation impacts: Web of
science, scopus, and dimensions as varying resonance chambers.
\emph{Scientometrics}, \emph{127}(5), 2413--2431.
\url{https://doi.org/10.1007/s11192-022-04309-6}

\bibitem[\citeproctext]{ref-Suber_2012}
Suber, P. (2012). \emph{Open access}. The MIT Press.
\url{https://doi.org/10.7551/mitpress/9286.001.0001}

\bibitem[\citeproctext]{ref-Visser_2021}
Visser, M., Eck, N. J. van, \& Waltman, L. (2021). Large-scale
comparison of bibliographic data sources: {Scopus, Web of Science,
Dimensions, Crossref, and Microsoft Academic}. \emph{Quantitative
Science Studies}, \emph{2}(1), 20--41.
\url{https://doi.org/10.1162/qss_a_00112}

\bibitem[\citeproctext]{ref-Zhang_2024}
Zhang, L., Cao, Z., Shang, Y., Sivertsen, G., \& Huang, Y. (2024).
Missing institutions in OpenAlex: Possible reasons, implications, and
solutions. \emph{Scientometrics}.
\url{https://doi.org/10.1007/s11192-023-04923-y}

\end{CSLReferences}

\end{document}