### 2025-04-15 analysis/manuscript.tex

\documentclass[a4paper,man,floatsintext,longtable,noextraspace,10pt]{apa6}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{hyperref}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
% make captions italic

% number lines
% \usepackage{lineno}
% \linenumbers

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
\begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
% allow citations to break across lines
\let\@cite@ofmt\@firstofone
% avoid brackets around text for \cite:
\def\@biblabel#1{}
\def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
{\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \setlength{\leftmargin}{\cslhangindent}
  \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
{\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% tightlist
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\shorttitle{}

% flow chart
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows, positioning, calc, shapes.multipart}

% Custom styles for the flowchart
\tikzstyle{datasource} = [rectangle, rounded corners, minimum width=2.5cm, minimum height=1cm,text centered, draw=blue!60, fill=blue!5]
\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=green!60, fill=green!5]
\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=red!60, fill=red!5]
\tikzstyle{arrow} = [thick,->,>=stealth]
\tikzstyle{output} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm, text centered, draw=purple!60, fill=purple!5]

\setlength\parindent{1.27cm}

\begin{document}
\thispagestyle{otherpage}

%\maketitle

% Change \title, \date, remove \author:
\begin{large}
\textbf{Estimating transformative agreement impact on hybrid open access: A comparative large-scale study using open metadata, Scopus and Web of Science}
\end{large}

% Add below \maketitle:
\newcommand{\orcid}{%
  \begingroup\normalfont
  \includegraphics[height=0.9em]{orcid_logo}% removed .png extension
  \endgroup
}

Najko Jahn\textsuperscript{1}\textsuperscript{*} 
(\orcid{} \href{https://orcid.org/0000-0001-5105-1463}{\color{black}{0000-0001-5105-1463}})

\textsuperscript{1} Göttingen State and University Library, University of Göttingen, Germany. \\

\textsuperscript{*} Correspondence: \href{mailto:najko.jahn@sub.uni-goettingen.de}{\color{black}{najko.jahn@sub.uni-goettingen.de}}

\section*{Abstract}
{}
{\textbf{Keywords}: hybrid open access, transformative agreements, scholarly publishing, big deals, bibliometrics}

\newpage

% QSS wants numbered sections
\setcounter{secnumdepth}{2}

\section{Introduction}\label{introduction}

Transformative agreements are a much-discussed library licensing model
for transitioning subscription-based journal publishing to full open
access. Although these agreements can vary considerably, they are mainly
aimed at hybrid journal bundles, allowing authors from participating
institutions to publish open access in these journals, while providing
reading access to the entire portfolio. Library consortia and funders
have begun coordinating how these agreements are negotiated, managed and
evaluated. At the same time, efforts are underway to improve
transparency through open data on contracts and payments. However, this
information is still only partially available, which limits large-scale
studies of the impact of transformative agreements on the transition of
hybrid journals to full open access. By comparing open metadata from
hoaddata, an openly available dataset on hybrid open access based on
metadata from Crossref, OpenAlex and the cOAlition S Journal Checker
Tool, with the established bibliometric databases Scopus and Web of
Science, this study aims to provide a robust estimate of the impact of
transformative agreements on hybrid open access. This comparative
approach informs bibliometric research and practice by highlighting the
strengths and weaknesses of each data source individually and
collectively.

Since their initial proposal by the Max Planck Digital Library (Schimmer
et al., 2015), transformative agreements have led to a rapid growth in
the number of negotiated agreements and open access articles (Jahn,
2025; McCabe \& Mueller-Langer, 2024; Rothfritz et al., 2024). Before,
only a small proportion of articles in hybrid journals were openly
available (Björk, 2012; Laakso \& Björk, 2016; Piwowar et al., 2018),
with publications from only a few European countries with dedicated
funding policies and the high-energy physics SCOAP\^{}3 consortium
standing out (Huang et al., 2020; Kohls \& Mele, 2018; Pinfield, 2015;
Robinson-Garcia et al., 2020). Related coordination activities, in
particular through the OA2020 and ESAC (Efficiency and Standards for
Article Charges) initiatives, have also led to improved workflows
between library consortia and publishers (Campbell et al., 2022;
Geschuhn \& Stone, 2017). However, critics have raised concerns about
these agreements, particularly regarding market concentration (Butler et
al., 2023; Shu \& Larivière, 2023), the failure of most hybrid journals
to convert to full open access (Kiley, 2024; Momeni et al., 2021),
reduced competition and incentives (McCabe \& Mueller-Langer, 2024;
Schmal, 2024), and continuation of publication fees that widen gaps
between well-resourced and under-resourced institutions (Babini et al.,
2022; Ross-Hellauer et al., 2022)

While consortia evaluations confirm the growth of open access through
transformative agreements and acknowledge increasing coordination, they
reach mixed policy conclusions. For instance, the British Jisc
evaluation (Brayman et al., 2024) found that the consortium's
transformative agreements had a significant impact on national open
access growth but a limited effect on the global open access transition,
suggesting a need to reevaluate their strategy. Similar policy
reconsiderations emerged in Norway and Sweden (Holden et al., 2023;
Widding, 2024). Meanwhile, the funder initiative cOAlition S recommended
ending financial support for these agreements in 2024, but still
considers hybrid open access to be compliant with funders' open access
policies. Conversely, the German DEAL consortium extended agreements
with major publishers until 2028, while also outside of Europe more and
more transformative agreements are being put in place (for instance in
Latin America, as described in Muñoz-Vélez et al. (2024)). Responding to
criticism, the Max Planck Digital Library recently outlined strategies
to evolve the original model, focusing on achieving full open access
while controlling costs and addressing identified weaknesses (Dér, 2025)

Despite controversies, transformative agreements can benefit the
landscape of open research information as library consortia negotiate
for comprehensive open metadata, helping avoid data analytics becoming a
commodity controlled by publishers (Aspesi \& Brand, 2020; McCabe \&
Mueller-Langer, 2024). For instance, ESAC recommends using Crossref
(Hendricks et al., 2020) to share metadata supporting workflows
including open access licenses and funding information (Geschuhn \&
Stone, 2017). Simultaneously, the bibliometric community pushes for more
open metadata through this publisher-driven DOI registration platform
and continuously monitors progress (N. J. van Eck \& Waltman, 2022).

Many open scholarly data services that build on Crossref benefit from
these developments. A prominent example is OpenAlex, which has gained
attention for its inclusivity, openness and comprehensive coverage of
publication metadata (Priem et al., 2022). In a recent evaluation,
Céspedes et al. (2025) reviewed the current evidence base on the
strengths and weaknesses of OpenAlex. They highlight both its increasing
use in bibliometric research and its adoption by research organisations.
Comparative studies have generally found OpenAlex to be a reliable
alternative to commercial databases (Alperin et al., 2024; Culbert et
al., 2024), although data quality issues occurred for key metadata
fields needed for open access analysis, including evidence of open
access status itself (Jahn et al., 2023; Simard et al., 2024), as well
as authorships (Culbert et al., 2024), affiliations (Zhang et al.,
2024), author roles (Jahn, 2025) and document types used to distinguish
between different article types (Haupka et al., 2024). However, these
studies highlight the rapid improvements and the new possibilities of
combining open metadata and the curation efforts of OpenAlex, in
conjunction with the growing awareness of the importance of open
metadata, as demonstrated by the Barcelona Declaration on Open Research
Information.

As a consequence, there is growing interest in using open bibliometric
databases such as Crossref and OpenAlex to assess the impact of
transformative agreements. Similar to established proprietary
bibliometric databases such as Scopus and Web of Science, these data
sources lack specific data points, particularly about open access
payments, which are typically not shared (Haustein et al., 2024).
However, bibliometric data can be enriched with open metadata about the
journals and institutions involved in transformative agreements (Jahn,
2025). Jonge et al. (2025) demonstrated this potential by combining data
from the cOAlition S Journal Checker tool and OpenAlex, achieving 89\%
accuracy when validated against Dutch research information. While open
data approaches show promise, understanding how these results compare
when the same methodology is applied to proprietary databases such as
Web of Science and Scopus remains unexplored.

This study compares open and proprietary data sources for evaluating
transformative agreements. It adapts the approach from Jahn (2025),
which combines data from the cOAlition S Journal Checker tool, Crossref
and OpenAlex, and applies the same methods to Web of Science and Scopus
data. This comparative approach allows the evaluation of different
bibliometric data sources for monitoring transformative agreements,
while assessing their respective strengths and limitations.
Specifically, this study compares the indexing of more than 13,000
hybrid journals across these sources, and examines whether consistent
results can be obtained in terms of open access adoption between 2019
and 2023, analysed by publisher and country. By comparing results across
these data sources, this large-scale study provides a more robust
understanding of transformative agreements' impact on the transition to
full open access. It also reveals the comparative strengths and
weaknesses of different bibliometric databases in measuring scholarly
publishing's shift toward openness

\section{Data and methods}\label{data-and-methods}

As shown in Figure \ref{fig:workflow}, the methodology involved
comparing hoaddata, an openly available collection of open research
information on hybrid open access, with the bibliometric databases Web
of Science and Scopus. This section introduces the initial data sources,
followed by a presentation of the necessary data processing steps to
obtain eligible articles enabled by transformative agreements using
author roles (first and corresponding) and harmonised affiliation data.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[node distance=2cm, thick,scale=1, every node/.style={scale=0.9}]    
% Data Sources
    \node (crossref) [datasource] {Crossref};
    \node (openAlex) [datasource, right=of crossref] {OpenAlex};
    \node (jct) [datasource, right=of openAlex] {JCT};
    \node (wos) [datasource, below=of crossref] {Web of Science};
    \node (scopus) [datasource, below=of jct] {Scopus};

    % Processing Steps
    \node (hoaddata) [datasource, below=1.5cm of $(crossref)!1!(openAlex)$] {hoaddata};
    \node (kb) [process, below=1cm of $(wos)!0.5!(scopus)$] {Journal and article retrieval (including open access)};

 % Matching Process - Fixed line break syntax
    \node (matching) [process, below=1.5cm of $(hoaddata)!1!(kb)$] {Harmonising author affiliations through ROR};

 % Aggregation
    \node (transmatch) [process, below=1.5cm of $(matching)!1!(matching)$] {Estimate article output under transformative agreements};

    % Output
    \node (analysis) [output, below=1.5cm of $(transmatch)!1!(transmatch)$] {Aggregated analytical datasets};

    % Arrows
    \draw [arrow] (crossref) -- (hoaddata);
    \draw [arrow] (openAlex) -- (hoaddata);
    \draw [arrow] (jct) -- (hoaddata);
    \draw [arrow] (hoaddata) -- (kb);
    \draw [arrow] (wos) -- (kb);
    \draw [arrow] (scopus) -- (kb);
    \draw [arrow] (kb) -- (matching);
    \draw [arrow] (matching) -- (transmatch);
    \draw [arrow] (transmatch) -- (analysis);

    % Legend
    \node [datasource, scale=0.7] at ($(analysis)+(4,0)$) {Data Sources};
    \node [process, scale=0.7] at ($(analysis)+(4,-1)$) {Processing Steps};
    \node [output, scale=0.7] at ($(analysis)+(4,-2)$) {Output};
\end{tikzpicture}
\caption{Data processing workflow for comparing hybrid open access uptake across bibliometric data sources. The workflow shows how data from different sources (hoaddata, derived from Crossref, OpenAlex, Transformative Agreement Data dump used by the cOAlition S Journal Checker Tool (JCT), the Web of Science, and Scopus) were processed to enable comparative analysis.}
\label{fig:workflow}
\end{figure}

\subsection{Data sources}\label{data-sources}

\subsubsection{hoaddata}\label{hoaddata}

hoaddata, developed and maintained by the author to support open access
monitoring and research (e.g. Jahn (2025)), is an R data package that
regularly collects information on hybrid open access uptake from
multiple openly available data sources (Jahn, 2024). It combines
article-level metadata from Crossref (Hendricks et al., 2020) and
affilation metadata from OpenAlex (Priem et al., 2022) with
transformative agreement information from the Transformative Agreement
Data dump used by the cOAlition S Journal Checker Tool (JCT)\footnote{\url{https://journalcheckertool.org/transformative-agreements/}},
which links journal and institutional data about participating research
organisations to agreements in the ESAC registry.

hoaddata follows good practices for computational reproducibility using
R (Marwick et al., 2018). The package, which includes data, code, a test
suite and documentation, is openly available on GitHub. To ensure
computational reproducibility while aggregating the data, a GitHub
Actions continuous integration and delivery (CI/CD) workflow handles
data retrieval from the SUB Göttingen's open scholarly data warehouse
based on Google BigQuery, which provides high-performant programmatic
access to monthly snapshots of Crossref and OpenAlex.\footnote{\url{https://subugoe.github.io/scholcomm_analytics/data.html}}
The workflow has run regularly to fetch updates from these data sources
since 2022. The package version used in this study is 0.3, containing
data from the Crossref 2024-08 dump provided to Crossref Metadata Plus
subscribers and the OpenAlex 2024-08-29 monthly dump. It covers
agreements collected between July 2021 to July 2024 from the JCT. This
version including the computation log is available on GitHub
(https://github.com/subugoe/hoaddata/releases/tag/v.0.3).

\subsubsection{Web of Science}\label{web-of-science}

Clarivate Analytics' Web of Science (WoS) is a well-established
proprietary bibliometric database consisting of several collections
(Birkle et al., 2020). Web of Science is a selective, base
research-focused database (Stahlschmidt \& Stephen, 2022; Visser et al.,
2021). The collections considered in this study were the Science
Citation Index Expanded (SCIE), the Social Sciences Citation Index
(SSCI) and the Arts \& Humanities Citation Index (AHCI).

These collections provides important data points for analysing open
access: author affiliations and roles, differentiation of journal
articles into document types representing different types of journal
contributions, such as original articles or reviews, and open access
status information derived from OurResearch's Unpaywall (Piwowar et al.,
2018), the same provider as OpenAlex. However, Web of Science lacks
information about journals and articles under transformative agreements.

For programmatic access to article-level data, this study used the
database of the Kompetenznetzwerk Bibliometrie (KB) in Germany. The KB
processes raw XML data provided by Clarivate Analytics, which is
ingested into an in-house PostgreSQL database under a uniform schema. To
support reproducibility, KB maintains annual snapshots of the database.
Accordingly, this study used the annual snapshot from April 2024
(wos\_b\_202404), which is considered to cover almost the entire
previous publication year (Schmidt et al., 2024).

\subsubsection{Scopus}\label{scopus}

Elsevier's Scopus, launched in 2004, is another widely used proprietary
bibliometric database for measuring research (Baas et al., 2020).
Similar to Web of Science, Scopus is selective with regard to the
journals it indexes. However, its journal coverage is much broader than
that of the Web of Science collections considered in this study, as it
also indexes a wider range of applied research journals (Singh et al.,
2021; Stahlschmidt \& Stephen, 2022; Visser et al., 2021). With detailed
metadata about article types, open access status information derived
from Unpaywall, author roles, and disambiguated affiliations, Scopus
also contains important data to assess open access uptake, although no
direct information regarding transformative agreements was available at
the time of the study.

This study used the Scopus annual snapshot of April 2024 as provided by
the KB (scp\_b\_202404). The same KB curation effort as for the Web of
Science was applied to the Scopus (Schmidt et al., 2024).

\subsection{Data processing steps}\label{data-processing-steps}

\subsubsection{Determining hybrid journal publication
volume}\label{determining-hybrid-journal-publication-volume}

Following Jahn (2025), the starting point was a unified dataset of
several safeguarded JCT snapshots\footnote{\url{https://github.com/njahn82/jct_data}}.
The JCT journal data were enriched with ISSN variants linked to an
ISSN-L. To identify hybrid journals, a comprehensive exclusion of fully
open access journals was performed using multiple journal lists
including the Directory of Open Access Journals (DOAJ). The resulting
hybrid journal data were made available via hoaddata and used to
determine the publication volume for each database independently.

hoaddata relies on Crossref for obtaining journal publication volume and
open access status through Creative Commons (CC) licence information
relative to the published version (``version of record''). The article
metadata included DOIs, publication dates, open access information as
well as author roles and affiliations. Publication years were determined
using the earliest known date of publication in a journal. In hoaddata,
this corresponded to Crossref's issued date. For Web of Science and
Scopus, the earliest publication date was used where available, with
Scopus dates specifically determined by the KB through version tracking
of the raw data.

Many transformative agreements typically cover only certain types of
journal articles, in particular original research articles including
reviews (Borrego et al., 2021). Because of limited information on these
document types in open scholarly data (Haupka et al., 2024), hoaddata
used an extended version of Unpaywall's paratext recognition approach to
exclude non-scholarly content (Piwowar et al., 2018). To exclude
conference supplements, which are also often not covered by
transformative agreements, only articles published in regular issues,
indicated by numerical pagination, were considered. For Web of Science
and Scopus, their established, mainly accurate document type
classifications (Donner, 2017; Maisano et al., 2025) were used to
identify original research articles and reviews, referred to as original
articles throughout this study.

\subsubsection{Identifying open access articles in hybrid
journals}\label{identifying-open-access-articles-in-hybrid-journals}

Articles in hybrid journals were considered open access when they were
made freely available under a CC license on publisher platforms. While
hoaddata obtained this information from Crossref license metadata, Web
of Science and Scopus relied on Unpaywall as evidence source. Unpaywall
also uses Crossref license metadata, but supplements them by parsing
publisher websites directly, addressing cases where publishers do not
provide machine-readable CC license information (Piwowar et al., 2018).
This additional parsing remains necessary despite transformative
agreement workflows recommend to deposit CC licenses during DOI
registration (Geschuhn \& Stone, 2017). Both Web of Science\footnote{\url{https://webofscience.help.clarivate.com/en-us/Content/open-access.html}}
and Scopus\footnote{\url{https://blog.scopus.com/posts/scopus-filters-for-open-access-type-and-green-oa-full-text-access-option}}
defined hybrid open access consistently as content available under CC
licenses on publisher platforms, distinguishing it from bronze open
access that lack such explicit license information, or use
publisher-specific licenses (Piwowar et al., 2018).

\subsubsection{Harmonising author affiliations across
databases}\label{harmonising-author-affiliations-across-databases}

Author affiliations were retrieved for both first and, if available,
corresponding authors to prepare the linking between articles and
institutions covered by transformative agreements. To improve data
retrieval, JCT institution data was enriched with ROR-IDs from
associated institutions, such as university hospitals or institutes of
large research organisations such as the Max Planck Society, according
to OpenAlex' institution entity. To handle different address variants,
database-specific affiliation identifiers were used: ROR-IDs from
OpenAlex for hoaddata, affiliation enhanced names for Web of Science,
and Scopus Affiliation Identifier. Additionally, ISO country codes were
retrieved for each author's address to compile country-level statistics.

Because neither Web of Science nor Scopus supported ROR-ID at the time
of the data retrieval, the institution identifier used by the JCT, a
two-step matching process was implemented to harmonise affiliation data.
First, 2,782,540 articles from 6,457 institutions with ROR-IDs in the
JCT data since 2017 (according to hoaddata) were processed to map first
authors' ROR-IDs to corresponding proprietary affiliation identifier in
Web of Science and Scopus using DOI matching. Then, an algorithm
selected the most frequent ROR ID and proprietary identifier pairs to
handle multiple affiliations and organisational hierarchy differences.

This process linked 6,375 ROR-IDs to 4,894 Scopus Affiliation IDs, and
6,034 ROR-IDs to 2,422 enhanced affiliation strings in Web of Science.
Quality evaluation through random sampling of 50 pairs revealed an error
rate of 22\% for Web of Science (11 mismatches) and 6\% for Scopus (3
mismatches). Upon inspection, these mismatches primarily occurred with
less-represented institutions having only a few publications, introduced
through multiple affiliations of single authors. The difference between
databases suggests that Scopus' affiliation control aligns more closely
with ROR than that of Web of Science.

\subsubsection{Estimating open access in hybrid journals covered by
transformative
agreements}\label{estimating-open-access-in-hybrid-journals-covered-by-transformative-agreements}

Based on these compiled matching tables, articles eligible under
transformative agreements could also be obtained from Web of Science and
Scopus, although they did not contain ROR-IDs used by the JCT. The
estimation of eligible articles followed Jahn (2025) and included a
matching of both journals and participating institutions according to
the JCT. The matching also took into account the duration of agreements
according to the ESAC registry, with only those matches where an
agreement was actually in place being considered for subsequent
analysis. A related study (Jonge et al., 2025), applied to publications
funded by the Dutch Research Council (NWO) and validated against
internal invoicing data, confirmed that such matching can accurately
identify most articles under transformative agreements.

\subsection{Data records}\label{data-records}

As a result of the comprehensive data processing described above,
datasets on open access in hybrid journals included in transformative
agreements were aggregated for each database at country and journal
level by year. Table~\ref{tbl-methods_overview_tab} provides a general
overview of the coverage between 2019 and 2023. It shows that the
majority of hybrid journals published at least one original research
article or review, collectively marked as ``original article'' in the
following, during the five-year period. These journals provided the
basis for the subsequent calculation of article-level indicators.

\begin{table}

\caption{\label{tbl-methods_overview_tab}Coverage of hybrid journals in
transformative agreements 2019-23.}

\centering{

\centering
\begin{tabular}[t]{lrrr}
\toprule
\textbf{} & \textbf{hoaddata*} & \textbf{Web of Science} & \textbf{Scopus}\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Hybrid journal coverage}}\\
\hspace{1em}Active journals & 12,890 & 8,655 & 11,888\\
\hspace{1em}Active journals (core) & 12,888 & 8,655 & 11,878\\
\hspace{1em}Active journals (core) with OA & 11,348 & 8,392 & 11,313\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Publication volume}}\\
\hspace{1em}Total published articles & 9,740,015 & 8,616,053 & 8,117,644\\
\hspace{1em}Original articles & 8,158,425 & 6,708,083 & 7,317,703\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Digital Object Identifier (DOI) coverage}}\\
\hspace{1em}Articles with DOI & 9,740,015 & 7,713,796 & 8,105,112\\
\hspace{1em}Original articles with DOI & 8,158,425 & 6,695,661 & 7,314,327\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Open Access (OA) metrics}}\\
\hspace{1em}OA articles & 998,699 & 1,112,758 & 974,099\\
\hspace{1em}Original OA articles & 969,817 & 1,019,784 & 922,578\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Original articles with affiliation data}}\\
\hspace{1em}First author articles & 7,242,542 & 6,294,855 & 7,232,017\\
\hspace{1em}Corresponding author articles & 5,534,207 & 6,291,441 & 6,898,487\\
\bottomrule
\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{*} Journal article metadata from Crossref, except affiliations from OpenAlex}\\
\end{tabular}

}

\end{table}%

While hoaddata only covers articles with a DOI, Scopus and Web of
Science publication indicators were calculated using their database
identifier. A subsequent comparison of DOI coverage shows that
non-original articles in Web of Science often lacked a DOI. This was
particularly the case for meeting abstracts, which are notably prevalent
in Health Sciences journals (Melero-Fuentes et al., 2025) and are not
indexed by Scopus (Donner, 2017). Open access indicators were aggregated
by DOI, because Unpaywall only collects information on open access
status for articles with a DOI. A closer look at original articles with
affiliation data, and in line with related research (Zhang et al.,
2024), reveals a lack of affiliation data in the case of OpenAlex, the
affiliation data source used by hoaddata, compared to Web of Science and
Scopus. In particular, only about two-thirds of the articles examined
provided corresponding author affiliations. For first authors, the
proportion was 89\%. At the time of writing, OpenAlex disclosed limited
coverage of corresponding authorship data \footnote{\url{https://docs.openalex.org/api-entities/works/work-object/authorship-object\#is_corresponding}}.
Therefore, only first author data for hoaddata were considered in the
following analysis.

\subsection{Data analysis}\label{data-analysis}

Throughout this largely automated data collection and analysis process,
Tidyverse tools (Wickham et al., 2019) for the R programming language (R
Core Team, 2024) were used. Rank correlations were calculated using the
Hmisc package (Harrell Jr, 2003). The source code analysis, including
all queries used to obtain the data, is available on GitHub (add link).
An interactive supplement exploring correlations between the data
sources examined by country and publisher is available on HuggingFace
Spaces: https://huggingface.co/spaces/najkoja/hoa\_replication.

\section{Results}\label{results}

This section first presents the indexing of hybrid open access by
comparing open data sources with the proprietary bibliometric databases
Scopus and Web of Science. Then, using the same methods, indicators at
the publisher and country level are calculated independently for each
database and compared with each other to assess the suitability of the
bibliometric databases for investigating transformative agreements.

\subsection{Coverage comparision}\label{coverage-comparision}

\subsubsection{Overview}\label{overview}

Figure \ref{fig-upset_coverage_results} presents the coverage of hybrid
journals included in transformative agreements, based on the
intersections of the examined data sources. The intersection sets of
journals and articles are visualised as an UpSet graph (Krassowski,
2020; Lex et al., 2014). The analysis included hybrid journals that
published at least one open access article between 2019 and 2023, based
on open access status information from each database. Only original
articles were considered in the analysis.

\begin{figure}[ht!]

\centering{

\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/fig-upset_coverage_results-1.pdf}

}

\caption{\label{fig-upset_coverage_results}Comparison of hybrid journal
indexing by data source, 2019-2023. Only hybrid journals present in the
cOAlition S Transformative Agreement Data dump with at least one open
access article are considered. A) presents the number of journals, B)
the number of articles (DOI), distinguishing between shared DOI corpus
and surplus in hoaddata. Box plots (C) shows the five-years journal
article volume (log-scale).}

\end{figure}%

Journal coverage analysis revealed that 66\% (n = 7,970) of hybrid
journals included in transformative agreements were indexed in all three
databases (Figure \ref{fig-upset_coverage_results}A). The second-largest
set consisted of journals indexed in both hoaddata and Scopus,
comprising 21\% (n = 2,595) of hybrid journals. Notably, 6\% (n = 739)
of journals were exclusively contained in hoaddata, while another 6\% (n
= 748) were only found in Scopus. Of these, 354 were also available in
Web of Science. Upon inspection, this group of hybrid journals
exclusively covered in the proprietary data sources mainly represented
hybrid journals for which no open access evidence could be retrieved
from Crossref, the open access evidence source for hoaddata.

In terms of article coverage, Figure \ref{fig-upset_coverage_results}B
shows the total publication volume per combination in terms of DOI
availability. The largest set of hybrid journals, covering all three
data sources, also represents the largest number of articles. In total,
these journals recorded 6,289,687 articles, represented by the blue bar.
They represented 94\% of original articles with a DOI indexed in the Web
of Science sample, and 86\% in Scopus. Another 657,697 articles were
exclusive to the intersection of Scopus and hoaddata. Exclusively in
hoaddata were 177,110 articles, and exclusively in the proprietary
databases were 325,194 articles.

Figure \ref{fig-upset_coverage_results}B also shows the surplus of
articles with DOI that were only available via hoaddata (grey area). In
the case of hybrid journals covered by all three data sources, 1,023,882
DOIs were only present in hoaddata. After validation at the DOI level
using the KB databases and manual inspection, the main reason for
missing DOI coverage in the proprietary database were insufficient
classification of journal content as original articles during the
compilation of hoaddata. Particularly, letters and editorials could not
be fully detected. Moreover, paratext recognition failed for 37\% of
DOIs to identify non-scholarly content such as front matters or reviewer
lists, which are generally not indexed by Scopus and Web of Science. To
a lesser extent, differences in publication and indexing dates were a
reason for non-overlapping DOIs.

Using overlapping DOIs, the publication volume between 2019 and 2023 was
also calculated for each journal. Figure
\ref{fig-upset_coverage_results}C illustrates the distribution for each
combination. It shows a large spread oj journal size across the journals
jointly covered by all three data sources. These journals published more
on average than journals covered by only one or two data sources, with
the exception of journals exclusively covered by Scopus and Web of
Science. In particular, journals covered exclusively by hoaddata were
substantially smaller. Upon inspection, these were often newly launched
hybrid journals, which explains the relatively low five-year publication
volume. An example is \emph{Digital society} that published 86 articles.
This hybrid journal was launched in 2022, being covered by various
Springer Nature transformative agreements since then.

\subsubsection{Coverage by publisher
portfolio}\label{coverage-by-publisher-portfolio}

Figure \ref{fig-upset_coverage_results_publisher} presents the coverage
of hybrid journals in transformative agreements across data sources from
2019 to 2023 with a focus on publisher portfolios. The analysis
highlights the dominance of the three largest publishers, Elsevier,
Springer Nature, and Wiley, which collectively accounted for 46\% of
hybrid journals and 62\% of articles published during this five-year
period. In terms of article volume, Elsevier led with 2,435,478 articles
(34\% of the total) published across 1,941 hybrid journals (16\% of the
total). Springer Nature followed with 1,237,700 articles (17\%) in 2,294
hybrid journals, although recording the largest number of hybrid
journals (19\%). Wiley accounted for 814,349 articles (11\%) in 1,337
hybrid journals (11\%). The remaining 54 publishers collectively
accounted for 2,733,994 articles (38\%) in 6,442 hybrid journals (54\%).

\begin{figure}[ht!]

\centering{

\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/fig-upset_coverage_results_publisher-1.pdf}

}

\caption{\label{fig-upset_coverage_results_publisher}Comparison of
hybrid journal indexing by data source and publisher, 2019-2023. Only
hybrid journals present in the cOAlition S Transformative Agreement Data
dump with at least one open access article are considered. A) presents
the percentage of journals by publisher, B) the number of articles by
publisher (shared DOI). Box plots (C) shows the five-years journal
article volume (log-scale) by publisher.}

\end{figure}%

The three largest publishers, Elsevier, Springer Nature, and Wiley, were
best represented in the intersection of all three data sources
(hoaddata, Scopus, and Web of Science). Together, they comprised 4,347
hybrid journals (55\% of the intersectional set) and dominated article
coverage (n = 4,143,663; 66\%), as determined through shared DOIs. When
examining publication volume per journal, (Figure
\ref{fig-upset_coverage_results_publisher}C), Elsevier published, on
average, the largest journals, followed by Springer Nature and Wiley.

Comparing publisher portfolios across different indexing sets
demonstrates that publishers were not represented uniformly. Notably,
Springer Nature exhibited 517 hybrid journals exclusively indexed in
both hoaddata and Scopus. This set included journals from the Chinese
Academy of Science, German-language medical journals, and Eastern
European publications including the \emph{Journal of Mathematical
Sciences}, which also publishes English-language translations of
Russian-language works. Additionally, this subset included titles with a
broader disciplinary focus such as \emph{SN Computer Science} and newly
launched hybrid journals like \emph{Nature Computational Science}, which
started in 2021 and was indexed in Scopus but not yet in Web of Science.
The set also captured discontinued journals, providing further insights
into the dynamics of journal publishing.

Examining publisher portfolios not covered by hoaddata but present in
Scopus or Web of Science identified several publishers with missing
Creative Commons (CC) license information in Crossref. In particular,
Emerald represented 321 journals with 86,269 articles, AIP Publishing
accounted for 24 journals with 64,898 articles, and World Scientific
recorded 87 journals and 42,531 articles. In total, 9 publishers did not
share CC licenses with Crossref and were therefore not represented in
hoaddata.

An inspection of individual journals also uncovered discrepancies in
Unpaywall's open access identification for certain publishers that
typically share CC license metadata with Crossref. Notably, some
subscription-only journals contained one or two articles erroneously
tagged as hybrid open access by Unpaywall, which were subsequently
reflected in Scopus and Web of Science. Examples of such
misclassifications include Elsevier's \emph{Journal of Bioscience and
Bioengineering} and Springer Nature's \emph{Journal of Mechanical
Science and Technology}.

\subsection{Open Access Indicator
Comparison}\label{open-access-indicator-comparison}

This section examines the uptake of open access in hybrid journals,
focusing on the influence of transformative agreements across hoaddata,
Scopus, and Web of Science. The aim was to assess whether consistent
results can be derived from these data sources despite differences in
coverage and methodologies. Following Jahn (2025), indicators were
calculated for each data source and comprise the number and proportion
of open access articles, including those enabled by transformative
agreements, from 2019 to 2023. For Web of Science and Scopus, the impact
of transformative agreements was estimated using both first and
corresponding authorships, while hoaddata indicator calculation was
limited to first author affiliations.

\begin{figure}[ht!]

\centering{

\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/fig-uptake_overview-1.pdf}

}

\caption{\label{fig-uptake_overview}Development of open access in hybrid
journals in transformative agreements by data source and author role,
2019-2023. Figure shows the open access percentage (A), the number (B)
and the percentage (C) of open access articles enabled by transformative
agreements.}

\end{figure}%

\subsubsection{Overview}\label{overview-1}

Figure \ref{fig-uptake_overview}A shows a moderate growth of open access
in hybrid journals, which is consistent across hoaddata (black line),
Scopus (yellow line), and Web of Science (blue line). According to
hoaddata, hybrid open access uptake increased from 6.4\% (n = 85,071) in
2019 to 19\% (n = 302,358) in 2023. Similarly, Scopus recorded a growth
from 6.5\% (n = 84,648) in 2019 to 19\% (n = 322,850) in 2023. However,
Web of Science recorded higher open access uptake in early years, before
converging to a similar level in 2023, from 11\% (n = 137,202) in 2019
to 20\% (n = 255,481) in 2023. This suggest a different approach towards
labelling hybrid open access by Web of Science.

Similarly, hybrid open access by transformative agreements substantially
increased between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and
C). Trends were consistent when measuring first (dashed line) and
corresponding author (dotted line) affiliations. According to Scopus,
479,297 open access articles could be attributed to transformative
agreements based on first author metadata (increasing from 15,341 to
206,084) and 489,262 using corresponding author metadata (from 15,444 to
210,816). Web of Science recorded 493,028 open access articles via
transformative agreements using first author metadata (increasing from
21,871 to 174,126) and 500,076 using corresponding author metadata (from
22,092 to 177,030). hoaddata, lacking corresponding author data, linked
501,649 articles to transformative agreements, increasing from 16,010 in
2019 to 183,757 in 2023.

From 2021 (hoadata, Scopus) resp. 2022 (Web of Science), transformative
agreements enabled the majority of open access articles in hybrid
journals. For first authors, the share ranged between 61\% (hoaddata),
64\% (Scopus), and 68 \% (Web of Science) in 2023. For corresponding
authors, the shares were slightly larger, with Scopus recording 65\% and
Web of Science 69\% in 2023. However, substantial hybrid open access was
still facilitated outside transformative agreements, likely through APCs
paid from discretionary research funds (Suber, 2012).

\subsubsection{Open access by
publishers}\label{open-access-by-publishers}

\begin{figure}[ht!]

\centering{

\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/fig-uptake_publisher-1.pdf}

}

\caption{\label{fig-uptake_publisher}Development of open access in
hybrid journals included in transformative agreements by data source,
author role and publisher, 2019-2023. Figure shows the open access
percentage (A), the number (B) and the percentage (C) of open access
articles enabled by transformative agreements.}

\end{figure}%

When considering open access trends by publisher (see Figure
\ref{fig-uptake_publisher}), the observed differences in early uptake
rates between hoaddata and Scopus compared to Web of Science can be
largely attributed to articles published in Elsevier hybrid journals,
the largest publisher in our sample. Both hoadata and Scopus reported a
steady increase in open access uptake between 2019 and 2023 (hoaddata
from 4\% to 13\%; Scopus from 5\% to 15\%). In contrast, Elsevier's
share remained relatively constant, increasing only slightly from 14\%
to 15\% according to Web of Science. Upon inspection, this discrepancy
is primarily due to articles in the publisher's open archive. These
articles, made freely available after an embargo period under Elsevier's
user license, were tagged as hybrid open access in Web of Science, even
though its documentation\footnote{\url{https://webofscience.help.clarivate.com/en-us/Content/open-access.html}}
specified that only articles under a CC license variant were considered.
Previous research (Haustein et al., 2024; Jahn et al., 2022) has shown
that Elsevier provided a substantial portion of its articles under this
license, explaining the relatively large and stable share of open access
over the years.

Differences in open access evidence are also evident for Wiley.
Specifically, Web of Science and Scopus recorded a drop in 2021 and 2022
compared to hoaddata. For these two years, hoaddata reported 32,445 more
open access articles than Scopus and 29,854 more than Web of Science.
This discrepancy is presumably due to challenges in fetching full-texts
by Unpaywall, the open access evidence source for Scopus and Web of
Science. According to Unpaywall's software version history, HTTP
redirects from Wiley's publisher platform prevented Unpaywall from
parsing license information from full-texts.\footnote{See Unpaywall
  version history related to Wiley fixes:
  \url{https://github.com/search?q=repo\%3Aourresearch\%2Foadoi+wiley&type=commits}}
hoaddata, which relies solely on Crossref metadata for open access
identification, was unaffected by these issues.

Despite these differences in open access evidence, the three data
sources show consistent temporal trends in hybrid open access enabled by
transformative agreements (see Figure \ref{fig-uptake_publisher}B and
C). Wiley emerged as the fastest-growing publisher in terms of open
access uptake, with more than 30\% of articles in hybrid journals
reported as open access in 2023 across the examined data sources,
followed by Springer Nature. Elsevier, recorded a later uptake, which is
consistent with the publisher's historical reluctance to engage in
negotiations with library consortia. However, by 2023, the share of open
access enabled by transformative agreements appeared to stabilise for
all three publishers (see Figure \ref{fig-uptake_publisher}C).
Interestingly, the differences between first and corresponding author
affiliations were more pronounced at the publisher level. In Scopus, for
example, the share of open access via transformative agreements measured
by corresponding authorship was higher for Springer Nature in 2023 than
when using first authorship.

\subsubsection{Open access by country}\label{open-access-by-country}

When comparing countries, consistent patterns were observed across data
sources for the five-year period 2019 to 2023. Figure
\ref{fig-uptake_country} presents hybrid open access indicators by
country, comparing hoaddata (x-axis) with Web of Science and Scopus
(y-axis). Indicators calculated from these proprietary databases are
shown for both first and corresponding authors, with full counting used
to account for multiple country affiliations (Hottenrott et al., 2021).

\begin{figure}[ht!]

\centering{

\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/fig-uptake_country-1.pdf}

}

\caption{\label{fig-uptake_country}Comparison of hybrid open access by
country, 2019-2023. Scatterplots distinguish between proprietary
databases Scopus and Web of Science, and author role. The x-axis shows
hoaddata indicators. A) Five-years article volume, B) open access
percentage in hybrid journals, (A and B limited to countries with
\textgreater{} 10.000 articles published), C) number and D) percentage
of open access articles enabled by transformative agreements (limited to
countries with \textgreater{} 1.000 open access articles enabled by
transformative agreements). Line represents line of equality. An
interactive version is accessible via:
\url{https://najkoja-hoa-replication.hf.space/}}

\end{figure}%

In terms of article output by country (see Figure
\ref{fig-uptake_country}A), strong positive correlations were observed
across the data sources and author roles (Spearman rank correlation
\(\rho > .9\), \(p < 0.001\)). Between 2019 and 2023, China was the most
productive country, followed by the United States and, by a certain
margin, India, the United Kingdom, and Germany. Analysis of authorship
roles revealed minimal variation, indicating that first and
corresponding authors were typically from the same country.

When examining the percentage of open access articles in hybrid journals
(see Figure \ref{fig-uptake_country}B), a different pattern emerged.
Authors affiliated with institutions from medium-sized European
countries, such as Sweden, the Netherlands, Finland, and Hungary,
provided a large proportion of their articles in open access. Germany
and the United Kingdom also had approximately 40\% of their output
available as open access. In contrast, non-OECD countries showed notably
lower adoption of hybrid open access, with South Africa being the only
BRICS member well-represented in the data. The United States also
demonstrated a relatively low proportion of open access articles. These
findings were consistent across all databases. However, France was
better represented in Web of Science, likely due to its agreement with
Elsevier starting in 2019, which allowed delayed open access under the
publisher's user license (Rabesandratana, 2019). This licence was not
classified as hybrid open access in either Scopus or hoaddata. In all
cases, Spearman rank correlations were \(\rho > .9\), \(p < 0.001\),
showing a high level of correlations between the databases and
authorship roles considered.

Transformative agreements appeared to be a key driver of national open
access growth (see Figure \ref{fig-uptake_country}C and D). OECD members
accounted for the majority of open access articles enabled by
transformative agreements. As a notable exception, South Africa again
featured prominently, as the South African National Library and
Information Consortium (SANLiC) successfully negotiated transformative
agreements with major publishers from 2022 onward.\footnote{\url{https://sanlic.ac.za/read-and-publish-agreements/}}
Results were consistent across data sources. However, Wiley's open
access surplus in hoaddata led to better rankings for countries where
Wiley played a substantial role, such as Germany, where the DEAL
consortium negotiated its first transformative agreements with publisher
that started in July 2019.

Medium-sized European countries again showed a high proportion of hybrid
open access through transformative agreements (see Figure
\ref{fig-uptake_country}D), highlighting the impact of this licensing
model across all three data sources. In contrast, the United States had
a low proportion of hybrid open access enabled by agreements, suggesting
that a substantial number of open access articles from US-based authors
were likely financed through other means.

In all cases, strong positive correlations were observed using
Spearman's rank correlation: \(\rho > .9\), \(p < 0.001\) between data
sources and authorship roles, when considering countries with a minimum
of 1,000 open access articles enabled by transformative agreements
between 2019 and 2023. When this limitation was removed, the correlation
remained strong (\(\rho > .85\), \(p < 0.001\)). This difference may
signal countries where only a few institutions had transformative
agreements in place, as opposed to those participating in national
consortia with broader participation.

\section{Discussion}\label{discussion}

This study of over 13,000 hybrid journals shows a substantial rise in
open access due to transformative agreements between 2019 and 2023,
although most articles remained paywalled. While transformative
agreements accounted for the majority of open access, many articles
continue to become open through the payment of individual publication
fees. Hybrid open access and transformative agreements remain
concentrated among a small group of large commercial publishers, with
high-income European countries---alongside South Africa---showing high
adoption rates. In contrast, the three most productive countries, China,
the United States, and India, showcase a substantially lower adoption in
transformative agreements. Open questions remain whether this uneven
distribution reflects temporary implementation gaps, inherent inequities
in the transformative agreement model, or a deliberate avoidance of such
agreements.

The findings were consistent across the investigated open data source
hoaddata, derived from Crossref and OpenAlex, and the established
proprietary bibliometric databases Scopus and Web of Science. Notably,
and aligning with previous studies (Akbaritabar et al., 2024; Alperin et
al., 2024) and rankings (N. J. van Eck et al., 2024). Overall, the
results show strong correlations by country affiliation, which supports
using open metadata for large-scale analyses of hybrid open access.
However, the observed differences in journal coverage and metadata
availability warrant further discussion in the following, affecting not
only open data sources but also proprietary databases when used in
isolation.

The coverage analysis reveals that hybrid journals are well indexed in
all three data sources, particularly in terms of article coverage,
reflecting the dominance of major publishers whose established journal
portfolios are comprehensively indexed in proprietary databases (Bellen
et al., 2024). Differences emerge for journals targeting practitioners
or local non-English language communities, with many such titles indexed
exclusively in Crossref and Scopus. Using Crossref as a bibliometric
database in hoaddata demonstrated particular strength in identifying
newly established hybrid journals, a notable finding given that
transformative agreements primarily target existing subscription-based
journals. The landscape of hybrid journal publishing thus differs
markedly from that of fully open access journals. Comparing the coverage
of OpenAlex, Scopus and Web of Science, Simard et al. (2024) indicate
that only half of the fully open access journals listed in the DOAJ are
also indexed in Scopus and Web of Science. Notably, journals that charge
no publication fees (``diamond journals'') are absent from the selective
Web of Science, which reinforces existing disparities in the indexing of
under-represented research communities and regions in selective
bibliometric databases (Simard et al., 2024).

A frequently reported limitation of studying open access with less
selective databases is the lack of corresponding authorship information
(Fraser et al., 2023; Haucap et al., 2021; Shu \& Larivière, 2023).
However, this analysis demonstrates that indicators based on first
authors, which have often been used as a proxy for determining open
access funding, and corresponding authors show a high level of
correlation, reflecting disciplinary norms in scholarly publishing with
regard to contributorship and author roles and positions (Larivière et
al., 2016). First authors typically conduct the main research underlying
a paper, while the corresponding author often supervises the research
(Fox et al., 2018; Mattsson et al., 2010). Unsurprisingly, measures
based on first or corresponding authorships are strongly correlated,
suggesting that these authors share the same country affiliation.
Moreover, in most cases, the first author is identical to the
corresponding author (Chinchilla-Rodríguez et al., 2024). Despite this
correlation, the study observed a recent slight increase in open access
articles by corresponding authors over first authors across Springer
Nature hybrid journals. This prompts questions about how institutional
open access sponsorship practices influence author roles and assignments
within co-author teams, especially as funding opportunities vary
(Gumpenberger et al., 2018). Previous research has highlighted how
institutionalised bibliometric practices can affect the valuation of
authorship positions (Helgesson, 2020), suggesting the need to monitor
potential influences of the availability of open access funding on
authorship roles (Maddi \& Silva, 2024).

Another critical data element in the study is affiliation data, which is
essential for estimating open access enabled by transformative
agreements. Although OpenAlex's affiliation coverage is less
comprehensive, which likely reduced the number of articles confidently
attributed to transformative agreements in this study, it still shows
high correlations with Scopus and Web of Science at the country level.
However, OpenAlex's native ROR-ID integration offers a distinct
advantage, allowing a more reliable identification of agreement-enabled
articles compared to Scopus and Web of Science, which require
reconciliation with proprietary organisation identifiers. Future studies
based on the Web of Science will benefit from the recent integration of
ROR-IDs, announced by the end of 2024.

The database comparison revealed important discrepancies in open access
evidence. Crucially, not all publishers share CC licence metadata via
Crossref, a limitation that becomes apparent when contrasting Crossref
with Unpaywall's data in Scopus and Web of Science. While Unpaywall can
detect such gaps by parsing publishers' websites for open access
licences, it nevertheless missed a substantial number of CC-licensed
open access articles from Wiley journals indexed in Crossref, likely due
to parsing errors on the publisher's website. This resulted in fewer
open access articles being recorded in Scopus and Web of Science.
Further inconsistencies emerged between Scopus and Web of Science,
despite both relying on Unpaywall: Web of Science erroneously labels
Elsevier's delayed open access as hybrid, whereas Scopus correctly
categorises it. Notably, Scopus and Crossref showed greater alignment
than observed in a related comparison between Scopus and OpenAlex
(Alperin et al., 2024). This likely reflects the use of a curated list
to identify hybrid open access journals, rather than relying solely on
article-level tags. Given these variations, which may compromise
comparability, research and monitoring exercises should avoid reliance
on a single source. Instead, the selection of open access evidence
should be cross-verified using multiple sources and snapshots that can
be used to track changes in the data over time (Huang et al., 2020).
Incorporating expert-curated journal lists can also help prevent
misclassification based on business models (Visser et al., 2021).

It is important to note that the study's estimates of articles from
institutions involved in transformative agreements are approximations
due to a lack of access to invoice data, which is not usually shared by
library consortia and publishers. Furthermore, the data sources about
transformative agreements including the cOAlition S Journal Checker Tool
and the ESAC registry are based on a voluntary effort, crowd-sourced
from various consortia. However, recent validation with Dutch research
information demonstrates the reliability of such an open approach for
the assessment of articles under transformative agreements (Jonge et
al., 2025). With enhanced open metadata compliance, attributable to
evolving standards and initiatives in support of negotiations with
publishers, particularly through the communities of the ESAC initiative
and Barcelona Declaration supporters, the situation is likely to
improve. In conclusion, this study has shown that open metadata are
well-suited for analysing the transition from subscription-based journal
publishing to full open access.

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Akbaritabar_2024}
Akbaritabar, A., Theile, T., \& Zagheni, E. (2024). Bilateral flows and
rates of international migration of scholars for 210 countries for the
period 1998-2020. \emph{Scientific Data}, \emph{11}(1).
\url{https://doi.org/10.1038/s41597-024-03655-9}

\bibitem[\citeproctext]{ref-alperin2024analysissuitabilityopenalexbibliometric}
Alperin, J. P., Portenoy, J., Demes, K., Larivière, V., \& Haustein, S.
(2024). An analysis of the suitability of OpenAlex for bibliometric
analyses. \emph{arXiv}. \url{https://arxiv.org/abs/2404.17663}

\bibitem[\citeproctext]{ref-Aspesi_2020}
Aspesi, C., \& Brand, A. (2020). In pursuit of open science, open access
is not enough. \emph{Science}, \emph{368}(6491), 574--577.
\url{https://doi.org/10.1126/science.aba3763}

\bibitem[\citeproctext]{ref-Baas_2020}
Baas, J., Schotten, M., Plume, A., Côté, G., \& Karimi, R. (2020).
Scopus as a curated, high-quality bibliometric data source for academic
research in quantitative science studies. \emph{Quantitative Science
Studies}, \emph{1}(1), 377--386.
\url{https://doi.org/10.1162/qss_a_00019}

\bibitem[\citeproctext]{ref-budapest}
Babini, D., Chan, L., Hagemann, M., Joseph, H., Kuchma, I., \& Suber, P.
(2022). \emph{{The Budapest Open Access Initiative-20th. Anniversary
recommendations (BOAI20)}}.
\url{https://www.budapestopenaccessinitiative.org/boai20/}

\bibitem[\citeproctext]{ref-vanbellen2024oligopolyacademicpublisherspersists}
Bellen, S. van, Alperin, J. P., \& Larivière, V. (2024). \emph{The
oligopoly of academic publishers persists in exclusive database}.
\url{https://arxiv.org/abs/2406.17893}

\bibitem[\citeproctext]{ref-Birkle_2020}
Birkle, C., Pendlebury, D. A., Schnell, J., \& Adams, J. (2020). {Web of
Science} as a data source for research on scientific and scholarly
activity. \emph{Quantitative Science Studies}, \emph{1}(1), 363--376.
\url{https://doi.org/10.1162/qss_a_00018}

\bibitem[\citeproctext]{ref-Bj_rk_2012}
Björk, B.-C. (2012). The hybrid model for open access publication of
scholarly articles: A failed experiment? \emph{Journal of the American
Society for Information Science and Technology}, \emph{63}(8),
1496--1504. \url{https://doi.org/10.1002/asi.22709}

\bibitem[\citeproctext]{ref-Borrego_2020}
Borrego, Á., Anglada, L., \& Abadal, E. (2021). Transformative
agreements: Do they pave the way to open access? \emph{Learned
Publishing}, \emph{34}(2), 216--232.
\url{https://doi.org/10.1002/leap.1347}

\bibitem[\citeproctext]{ref-Jisc_2024}
Brayman, K., Devenney, A., Dobson, H., Marques, M., \& Vernon, A.
(2024). \emph{A review of transitional agreements in the {UK}}. Zenodo.
\url{https://doi.org/10.5281/zenodo.10787392}

\bibitem[\citeproctext]{ref-Butler_2023}
Butler, L.-A., Matthias, L., Simard, M.-A., Mongeon, P., \& Haustein, S.
(2023). The oligopoly's shift to open access: How the big five academic
publishers profit from article processing charges. \emph{Quantitative
Science Studies}, \emph{4}(4), 778--799.
\url{https://doi.org/10.1162/qss_a_00272}

\bibitem[\citeproctext]{ref-ifla}
Campbell, C., Dér, Á., Geschuhn, K., \& Valente, A. (2022). How are
transformative agreements transforming libraries? \emph{87th IFLA World
Library and Information Congress (WLIC) / 2022 in Dublin, Ireland}.
\url{https://repository.ifla.org/handle/123456789/1973}

\bibitem[\citeproctext]{ref-C_spedes_2025}
Céspedes, L., Kozlowski, D., Pradier, C., Sainte‐Marie, M. H., Shokida,
N. S., Benz, P., Poitras, C., Ninkov, A. B., Ebrahimy, S., Ayeni, P.,
Filali, S., Li, B., \& Larivière, V. (2025). Evaluating the linguistic
coverage of {OpenAlex}: An assessment of metadata accuracy and
completeness. \emph{Journal of the Association for Information Science
and Technology}. \url{https://doi.org/10.1002/asi.24979}

\bibitem[\citeproctext]{ref-Chinchilla_Rodr_guez_2024}
Chinchilla-Rodríguez, Z., Costas, R., Robinson-García, N., \& Larivière,
V. (2024). Examining the quality of the corresponding authorship field
in {Web of Science} and {Scopus}. \emph{Quantitative Science Studies},
\emph{5}(1), 76--97. \url{https://doi.org/10.1162/qss_a_00288}

\bibitem[\citeproctext]{ref-culbert2024referencecoverageanalysisopenalex}
Culbert, J., Hobert, A., Jahn, N., Haupka, N., Schmidt, M., Donner, P.,
\& Mayr, P. (2024). \emph{Reference coverage analysis of OpenAlex
compared to web of science and scopus}.
\url{https://arxiv.org/abs/2401.16359}

\bibitem[\citeproctext]{ref-D_r_2025}
Dér, Á. (2025). What gets missed in the discourse on transformative
agreements. \emph{Katina Magazine}.
\url{https://doi.org/10.1146/katina-20250212-1}

\bibitem[\citeproctext]{ref-Donner_2017}
Donner, P. (2017). Document type assignment accuracy in the journal
citation index data of {Web of Science}. \emph{Scientometrics},
\emph{113}(1), 219--236. \url{https://doi.org/10.1007/s11192-017-2483-y}

\bibitem[\citeproctext]{ref-eck_crossref_2022}
Eck, N. J. van, \& Waltman, L. (2022). \emph{Crossref as a source of
open bibliographic metadata}. MetaArXiv.
\url{https://doi.org/10.31222/osf.io/smxe5}

\bibitem[\citeproctext]{ref-eck_2024_6bx73}
Eck, N. J. van, Waltman, L., \& Neijssel, M. (2024). \emph{{Launch of
the CWTS Leiden Ranking Open Edition 2024}}. Leiden Madtrics.
\url{https://doi.org/10.59350/r512t-r8h93}

\bibitem[\citeproctext]{ref-Fox_2018}
Fox, C. W., Ritchey, J. P., \& Paine, C. E. T. (2018). Patterns of
authorship in ecology and evolution: First, last, and corresponding
authorship vary with gender and geography. \emph{Ecology and Evolution},
\emph{8}(23), 11492--11507. \url{https://doi.org/10.1002/ece3.4584}

\bibitem[\citeproctext]{ref-Fraser_2023}
Fraser, N., Hobert, A., Jahn, N., Mayr, P., \& Peters, I. (2023). No
deal: German researchers' publishing and citing behaviors after big deal
negotiations with {Elsevier}. \emph{Quantitative Science Studies},
\emph{4}(2), 325--352. \url{https://doi.org/10.1162/qss_a_00255}

\bibitem[\citeproctext]{ref-Geschuhn_2017}
Geschuhn, K., \& Stone, G. (2017). It's the workflows, stupid! What is
required to make {``offsetting''} work for the open access transition.
\emph{Insights the {UKSG} Journal}, \emph{30}(3), 103--114.
\url{https://doi.org/10.1629/uksg.391}

\bibitem[\citeproctext]{ref-Gumpenberger_2018}
Gumpenberger, C., Hölbling, L., \& Gorraiz, J. I. (2018). On the issues
of a {``corresponding author''} field-based monitoring approach for gold
open access publications and derivative cost calculations.
\emph{Frontiers in Research Metrics and Analytics}, \emph{3}.
\url{https://doi.org/10.3389/frma.2018.00001}

\bibitem[\citeproctext]{ref-Harrell_Jr_2003}
Harrell Jr, F. E. (2003). Hmisc: Harrell miscellaneous. In \emph{CRAN:
Contributed Packages}. The R Foundation.
\url{https://doi.org/10.32614/cran.package.hmisc}

\bibitem[\citeproctext]{ref-Haucap_2021}
Haucap, J., Moshgbar, N., \& Schmal, W. B. (2021). The impact of the
{German {``DEAL''}} on competition in the academic publishing market.
\emph{Managerial and Decision Economics}, \emph{42}(8), 2027--2049.
\url{https://doi.org/10.1002/mde.3493}

\bibitem[\citeproctext]{ref-Haupka_2024}
Haupka, N., Culbert, J. H., Schniedermann, A., Jahn, N., \& Mayr, P.
(2024). \emph{Analysis of the publication and document types in
OpenAlex, web of science, scopus, pubmed and semantic scholar}. arXiv.
\url{https://doi.org/10.48550/ARXIV.2406.15154}

\bibitem[\citeproctext]{ref-Haustein_APC_2024}
Haustein, S., Schares, E., Alperin, J. P., Hare, M., Butler, L.-A., \&
Schönfelder, N. (2024). \emph{Estimating global article processing
charges paid to six publishers for open access between 2019 and 2023}.
\url{https://arxiv.org/abs/2407.16551}

\bibitem[\citeproctext]{ref-Helgesson_2020}
Helgesson, G. (2020). Authorship order and effects of changing
bibliometrics practices. \emph{Research Ethics}, \emph{16}(1--2), 1--7.
\url{https://doi.org/10.1177/1747016119898403}

\bibitem[\citeproctext]{ref-Hendricks_2020}
Hendricks, G., Tkaczyk, D., Lin, J., \& Feeney, P. (2020). Crossref: The
sustainable source of community-owned scholarly metadata.
\emph{Quantitative Science Studies}, \emph{1}(1), 414--427.
\url{https://doi.org/10.1162/qss_a_00022}

\bibitem[\citeproctext]{ref-norway}
Holden, L., Skoie, M., Røeggen, V., Bjerde, K. W., Wenaas, L., Bakke,
P., Løvhaug, J. W., Karlsen, E. S., \& Qvenild, M. (2023).
\emph{Strategi for vitenskapelig publisering etter 2024}. Sikt.
\url{https://doi.org/10.18711/2KZ1-BA97}

\bibitem[\citeproctext]{ref-Hottenrott_2021}
Hottenrott, H., Rose, M. E., \& Lawson, C. (2021). The rise of multiple
institutional affiliations in academia. \emph{Journal of the Association
for Information Science and Technology}, \emph{72}(8), 1039--1058.
\url{https://doi.org/10.1002/asi.24472}

\bibitem[\citeproctext]{ref-Huang_2020}
Huang, C.-K., Neylon, C., Hosking, R., Montgomery, L., Wilson, K.,
Ozaygen, A., \& Brookes-Kenworthy, C. (2020). \emph{Evaluating
institutional open access performance: Sensitivity analysis}. bioRxiv.
\url{https://doi.org/10.1101/2020.03.19.998542}

\bibitem[\citeproctext]{ref-hoaddata}
Jahn, N. (2024). \emph{Hoaddata: Data about hybrid open access journal
publishing}. \url{https://github.com/subugoe/hoaddata}

\bibitem[\citeproctext]{ref-Jahn_2025}
Jahn, N. (2025). How open are hybrid journals included in transformative
agreements? \emph{Quantitative Science Studies}, \emph{6}, 242--262.
\url{https://doi.org/10.1162/qss_a_00348}

\bibitem[\citeproctext]{ref-jahn2023analysing}
Jahn, N., Haupka, N., \& Hobert, A. (2023). \emph{Analysing and
reclassifying open access information in OpenAlex}. {Scholarly
Communication Analytics Blog}.
\url{https://subugoe.github.io/scholcomm_analytics/posts/oalex_oa_status/}

\bibitem[\citeproctext]{ref-Jahn_2021}
Jahn, N., Matthias, L., \& Laakso, M. (2022). Toward transparency of
hybrid open access through publisher-provided metadata: An article-level
study of {Elsevier}. \emph{Journal of the Association for Information
Science and Technology}, \emph{73}(1), 104--118.
\url{https://doi.org/10.1002/asi.24549}

\bibitem[\citeproctext]{ref-de_Jonge_2025}
Jonge, H. de, Kramer, B., \& Sondervan, J. (2025). \emph{Tracking
transformative agreements through open metadata: Method and validation
using {Dutch Research Council NWO} funded papers}. MetaArXiv.
\url{https://doi.org/10.31222/osf.io/tz6be_v1}

\bibitem[\citeproctext]{ref-Kiley_2024}
Kiley, R. (2024). \emph{{Transformative Journals}: Analysis from the
2023 reports}. {cOAlition S}.
\url{https://www.coalition-s.org/blog/transformative-journals-analysis-from-the-2023-reports/}

\bibitem[\citeproctext]{ref-Kohls_2018}
Kohls, A., \& Mele, S. (2018). Converting the literature of a scientific
field to open access through global collaboration: The experience of
SCOAP3 in particle physics. \emph{Publications}, \emph{6}(2), 15.
\url{https://doi.org/10.3390/publications6020015}

\bibitem[\citeproctext]{ref-ComplexUpset}
Krassowski, M. (2020). \emph{ComplexUpset}.
\url{https://doi.org/10.5281/zenodo.3700590}

\bibitem[\citeproctext]{ref-Laakso_2016}
Laakso, M., \& Björk, B.-C. (2016). Hybrid open access--a longitudinal
study. \emph{Journal of Informetrics}, \emph{10}(4), 919--932.
\url{https://doi.org/10.1016/j.joi.2016.08.002}

\bibitem[\citeproctext]{ref-Larivi_re_2016}
Larivière, V., Desrochers, N., Macaluso, B., Mongeon, P., Paul-Hus, A.,
\& Sugimoto, C. R. (2016). Contributorship and division of labor in
knowledge production. \emph{Social Studies of Science}, \emph{46}(3),
417--435. \url{https://doi.org/10.1177/0306312716650046}

\bibitem[\citeproctext]{ref-Lex_2014}
Lex, A., Gehlenborg, N., Strobelt, H., Vuillemot, R., \& Pfister, H.
(2014). UpSet: Visualization of intersecting sets. \emph{IEEE
Transactions on Visualization and Computer Graphics}, \emph{20}(12),
1983--1992. \url{https://doi.org/10.1109/tvcg.2014.2346248}

\bibitem[\citeproctext]{ref-Maddi_2024}
Maddi, A., \& Silva, J. A. T. da. (2024). Beyond authorship: Analyzing
contributions in PLOS ONE and the challenges of appropriate attribution.
\emph{Journal of Data and Information Science}, \emph{9}(3), 88--115.
\url{https://doi.org/10.2478/jdis-2024-0015}

\bibitem[\citeproctext]{ref-Maisano_2025}
Maisano, D. A., Mastrogiacomo, L., Ferrara, L., \& Franceschini, F.
(2025). A large-scale semi-automated approach for assessing
document-type classification errors in bibliometric databases.
\emph{Scientometrics}. \url{https://doi.org/10.1007/s11192-025-05244-y}

\bibitem[\citeproctext]{ref-Marwick_2018}
Marwick, B., Boettiger, C., \& Mullen, L. (2018). Packaging data
analytical work reproducibly using {R} (and friends). \emph{The American
Statistician}, \emph{72}(1), 80--88.
\url{https://doi.org/10.1080/00031305.2017.1375986}

\bibitem[\citeproctext]{ref-Mattsson_2010}
Mattsson, P., Sundberg, C. J., \& Laget, P. (2010). Is correspondence
reflected in the author position? A bibliometric study of the relation
between corresponding author and byline position. \emph{Scientometrics},
\emph{87}(1), 99--105. \url{https://doi.org/10.1007/s11192-010-0310-9}

\bibitem[\citeproctext]{ref-McCabe_2024}
McCabe, M. J., \& Mueller-Langer, F. (2024). Open access is shaping
scientific communication. \emph{Science}, \emph{385}(6714), 1170--1172.
\url{https://doi.org/10.1126/science.adp8882}

\bibitem[\citeproctext]{ref-Melero_Fuentes_2025}
Melero-Fuentes, D., Aguilar-Moya, R., Valderrama-Zurián, J.-C., \&
Gorraiz, J. (2025). Evolution and effect of meeting abstracts in JCR
journals. \emph{Journal of Informetrics}, \emph{19}(1), 101631.
\url{https://doi.org/10.1016/j.joi.2024.101631}

\bibitem[\citeproctext]{ref-Momeni_2021}
Momeni, F., Mayr, P., Fraser, N., \& Peters, I. (2021). What happens
when a journal converts to open access? A bibliometric analysis.
\emph{Scientometrics}, \emph{126}(12), 9811--9827.
\url{https://doi.org/10.1007/s11192-021-03972-5}

\bibitem[\citeproctext]{ref-Mu_oz_V_lez_2024}
Muñoz-Vélez, H., Pallares, C., Echavarría, A. F., Contreras, J., Pavas,
A., Bello, D., Rendón, C., Calderón-Rojas, J., \& Garzón, F. (2024).
Strategies for negotiating and signing transformative agreements in the
{Global South}: The {Colombia Consortium} experience. \emph{Journal of
Library Administration}, \emph{64}(1), 80--98.
\url{https://doi.org/10.1080/01930826.2023.2287945}

\bibitem[\citeproctext]{ref-Pinfield_2015}
Pinfield, S. (2015). Making open access work: The {``state-of-the-art''}
in providing open access to scholarly literature. \emph{Online
Information Review}, \emph{39}(5), 604--636.
\url{https://doi.org/10.1108/oir-05-2015-0167}

\bibitem[\citeproctext]{ref-Piwowar_2018}
Piwowar, H., Priem, J., Larivière, V., Alperin, J. P., Matthias, L.,
Norlander, B., Farley, A., West, J., \& Haustein, S. (2018). The state
of {OA}: A large-scale analysis of the prevalence and impact of open
access articles. \emph{{PeerJ}}, \emph{6}, e4375.
\url{https://doi.org/10.7717/peerj.4375}

\bibitem[\citeproctext]{ref-priem2022openalex}
Priem, J., Piwowar, H., \& Orr, R. (2022). \emph{OpenAlex: A fully-open
index of scholarly works, authors, venues, institutions, and concepts}.
\url{https://arxiv.org/abs/2205.01833}

\bibitem[\citeproctext]{ref-r}
R Core Team. (2024). \emph{R: A language and environment for statistical
computing}. R Foundation for Statistical Computing.
\url{https://www.R-project.org/}

\bibitem[\citeproctext]{ref-Rabesandratana_2019}
Rabesandratana, T. (2019). Elsevier deal with {France} disappoints
open-access advocates. \emph{Science}.
\url{https://doi.org/10.1126/science.aba5656}

\bibitem[\citeproctext]{ref-Robinson_Garcia_2020}
Robinson-Garcia, N., Costas, R., \& Leeuwen, T. N. van. (2020). Open
access uptake by universities worldwide. \emph{{PeerJ}}, \emph{8},
e9410. \url{https://doi.org/10.7717/peerj.9410}

\bibitem[\citeproctext]{ref-Ross_Hellauer_2022}
Ross-Hellauer, T., Reichmann, S., Cole, N. L., Fessl, A., Klebel, T., \&
Pontika, N. (2022). Dynamics of cumulative advantage and threats to
equity in open science: A scoping review. \emph{Royal Society Open
Science}, \emph{9}(1). \url{https://doi.org/10.1098/rsos.211032}

\bibitem[\citeproctext]{ref-Rothfritz_2024}
Rothfritz, L., Schmal, W. B., \& Herb, U. (2024). \emph{Trapped in
transformative agreements? A multifaceted analysis of \textgreater1,000
contracts}. \url{https://arxiv.org/abs/2409.20224}

\bibitem[\citeproctext]{ref-Schimmer_2015}
Schimmer, R., Geschuhn, K., \& Vogler, A. (2015). \emph{{Disrupting the
subscription journals'business model for the necessary large-scale
transformation to open access}}. Max Planck Digital Library.
\url{https://doi.org/10.17617/1.3}

\bibitem[\citeproctext]{ref-Schmal_competition}
Schmal, W. B. (2024). \emph{La révolution dévore ses enfants: Pricing
implications of transformative agreements}. arXiv.
\url{https://doi.org/10.48550/ARXIV.2403.03597}

\bibitem[\citeproctext]{ref-schmidt_2024_13935407}
Schmidt, M., Rimmert, C., Stephen, D., Lenke, C., Donner, P., Gärtner,
S., Taubert, N., Bausenwein, T., \& Stahlschmidt, S. (2024). \emph{The
data infrastructure of the {German Kompetenznetzwerk Bibliometrie}: An
enabling intermediary between raw data and analysis}. Zenodo.
\url{https://doi.org/10.5281/zenodo.13935407}

\bibitem[\citeproctext]{ref-Shu_2023}
Shu, F., \& Larivière, V. (2023). The oligopoly of open access
publishing. \emph{Scientometrics}, \emph{129}(1), 519--536.
\url{https://doi.org/10.1007/s11192-023-04876-2}

\bibitem[\citeproctext]{ref-Simard_2024}
Simard, M.-A., Basson, I., Hare, M., Lariviere, V., \& Mongeon, P.
(2024). \emph{The open access coverage of {OpenAlex}, {Scopus} and {Web
of Science}}. \url{https://doi.org/10.48550/ARXIV.2404.01985}

\bibitem[\citeproctext]{ref-Singh_2021}
Singh, V. K., Singh, P., Karmakar, M., Leta, J., \& Mayr, P. (2021). The
journal coverage of web of science, scopus and dimensions: A comparative
analysis. \emph{Scientometrics}, \emph{126}(6), 5113--5142.
\url{https://doi.org/10.1007/s11192-021-03948-5}

\bibitem[\citeproctext]{ref-Stahlschmidt_2022}
Stahlschmidt, S., \& Stephen, D. (2022). From indexation policies
through citation networks to normalized citation impacts: {Web of
Science}, {Scopus}, and {Dimensions} as varying resonance chambers.
\emph{Scientometrics}, \emph{127}(5), 2413--2431.
\url{https://doi.org/10.1007/s11192-022-04309-6}

\bibitem[\citeproctext]{ref-Suber_2012}
Suber, P. (2012). \emph{Open access}. The MIT Press.
\url{https://doi.org/10.7551/mitpress/9286.001.0001}

\bibitem[\citeproctext]{ref-Visser_2021}
Visser, M., Eck, N. J. van, \& Waltman, L. (2021). Large-scale
comparison of bibliographic data sources: {Scopus, Web of Science,
Dimensions, Crossref, and Microsoft Academic}. \emph{Quantitative
Science Studies}, \emph{2}(1), 20--41.
\url{https://doi.org/10.1162/qss_a_00112}

\bibitem[\citeproctext]{ref-tidyverse}
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M.,
Pedersen, T. L., Miller, E., Bache, S. M., Müller, K., Ooms, J.,
Robinson, D., Seidel, D. P., Spinu, V., \ldots{} Yutani, H. (2019).
Welcome to the {tidyverse}. \emph{Journal of Open Source Software},
\emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}

\bibitem[\citeproctext]{ref-Widding_2024}
Widding, A. S. (2024). Beyond transformative agreements: Ways forward
for universities. \emph{European Review}, \emph{32}(S1), S28--S38.
\url{https://doi.org/10.1017/s1062798724000036}

\bibitem[\citeproctext]{ref-Zhang_2024}
Zhang, L., Cao, Z., Shang, Y., Sivertsen, G., \& Huang, Y. (2024).
Missing institutions in OpenAlex: Possible reasons, implications, and
solutions. \emph{Scientometrics}.
\url{https://doi.org/10.1007/s11192-023-04923-y}

\end{CSLReferences}

\end{document}